---
title: "2b_BBS_stacked_single-season occupancy_CONI"
author: "Marie I. Tosa"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load packages

require(unmarked)
require(plyr)
require(dplyr)
require(tidyr)
require(chron) #for times function

require(ggplot2)
require(ggpubr)
```


## Create functions to plot figures
```{r}
vals.coefs <- function(mod)
{
coefs <- coef(mod) %>% tibble::enframe() #extract coefficients for covariates
se <- SE(mod) %>% tibble::enframe() #extract SEs for covariates

ci.psi <- confint(mod, type="state") %>% tibble::enframe()
ci.det <- confint(mod, type="det") %>% tibble::enframe()
ci <- rbind(ci.psi, ci.det) #combine psi and det CIs
ci <- data.frame(name=ci$name, lower=ci$value[,1], upper=ci$value[,2]) #need to split up ci so can read as data frame with lower and upper bounds

coefs <- merge(coefs, se, by="name") #merge coefficient values and ses
coefs <- merge(coefs, ci) #merge coefs with CIs
coefs <- data.frame(coefs)
names(coefs) <- c("par", "est", "se", "lower","upper")

coefs$p <- gsub(substr(coefs$par, 0,3), pattern="p\\([A-z]", replacement="det")
coefs$name <- gsub(coefs$par, pattern="scale\\(", replacement="")
coefs$name <- gsub(coefs$name, pattern="p\\(", replacement="")
coefs$name <- gsub(coefs$name, pattern="psi\\(", replacement="")
coefs$name <- gsub(coefs$name, pattern="\\)\\)", replacement="")
coefs$name <- gsub(coefs$name, pattern="Int\\)", replacement="Int")
coefs$name <- gsub(coefs$name, pattern="\\)", replacement="")
coefs$name <- gsub(coefs$name, pattern="I\\(", replacement="")
return(coefs)
}

ylabel <- data.frame(type=c("det","state"), label=c("detection","occupancy"))
plot.marginal <- function(sp, mod, par, par.name, type)
{
  nd.og <- nd.mean[,names(nd.mean) != par] %>% slice(rep(1:n(), times=1000))
  nd <- data.frame(nd.og,
                   par=seq(min(site.stacked[,par]), max(site.stacked[,par]), length.out=1000))
  
  names(nd) <- gsub(names(nd), pattern="par", replacement=par)
  marginal <- predict(mod, type=type, newdata=nd)
  
  nd$StartTime <- times(nd$StartTime) #convert start time back to character

  o <- ggplot(marginal, aes(x=nd[,par], y=Predicted)) +
    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.5, fill="grey75") +
    geom_rug(data=site.stacked, aes(x=site.stacked[,par], y=0)) +
    ylim(0,1) +
    xlab(par.name) + ylab(paste("Predicted probability of", ylabel[ylabel$type == type,]$label)) +
    geom_line() + theme_bw(base_size=20) + theme(panel.grid=element_blank())
  print(o)
  ggsave(o, filename=paste("../Figures/occupancy_BBS/", sp, "_stacked_occupancy_",type,"_", par,".tiff", sep=""), height=6, width=6, unit="in", dpi=300, compression="lzw")
  return(o)
}

```

## load data
Created detection history data files for each species in 2a_BBS_nightjar_detection_timing.Rmd

```{r}
sp <- "coni"

data <- read.table(paste("../Data_processed/BBS_2022/BBS_", sp, ".txt", sep=""), sep=",", header=T)
summary(data)

table(data$Year)
table(data$Month)
table(data[,c("StopTotal","Year")])

#fix start time
data[data$StartTime == 0,]$StartTime <- NA
data$StartTime <- times(paste(substr(data$StartTime, 0,1), ":", substr(data$StartTime, 2,3), ":00", sep=""))

data$StartTemp.F <- as.numeric(data$StartTemp.F)
data$EndTemp.F <- as.numeric(data$EndTemp.F)

#remove rows with incomplete data
data <- data[!is.na(data$StartTemp.F),]
# data <- data[!is.na(data$EndTemp.F),]

data <- data[!is.na(data$StartTime),]

#remove years before 1997 since hardly any detections then
data <- data[data$Year >= 1997,]

#####
data$StateAbb <- factor(data$StateAbb)
data$Month <- factor(data$Month)

#MA 6/9/1991 temp is 11-19 F?
#ME 6/21/1993 temp is 17-17F?

```

## create umf for stacked single-season occupancy model
```{r}
sp.stacked <- data[,grep(names(data), pattern="Stop[0-9]")]
sp.stacked[sp.stacked > 1 & !is.na(sp.stacked)] <- 1 #convert > 1 to 1 (presence/absence instead of counts)

#check detection histories
unique(sp.stacked)

colSums(sp.stacked) #can we cut off detections anywhere? aka reduce the number of stops
plot(colSums(sp.stacked), cex=2, col="black", pch=20)

sp.stacked <- sp.stacked[,1:10] #only take the first 10 stops since others are surveys mid-day

######
#siteCovs
site.stacked <- data[,!names(data) %in% grep(names(data), pattern="Stop", value=T)] #siteCovs

#add in transformations for variables, if any
#
#
site.stacked$StartTime2 <- scale(site.stacked$StartTime)^2
site.stacked$jday2 <- scale(site.stacked$jday)^2


######
#obsCovs
stop.df <- matrix(rep(1:length(names(sp.stacked)), times=nrow(sp.stacked)), byrow=T, ncol=length(names(sp.stacked))) #add obsCovs, dimensions the same as y data
# stop2.df <- stop.df^2

umf <- unmarkedFrameOccu(y=sp.stacked, siteCovs=site.stacked, obsCovs=list(stop=stop.df))
head(umf)

# tiff(filename=paste("../Figures/occupancy_BBS/",sp,"dethist_BBS_stacked.tiff", sep=""), height=10, width=8, units="in", compression="lzw", res=400)
# plot(umf)
# dev.off()

plot(umf)
summary(umf)
```


## run detection models for stacked single season occupancy models
stops = 15
31942 sites
Maximum number of observations per site: 15 
Mean number of observations per site: 15 
Sites with at least one detection: 1451 

Tabulation of y observations:
     0      1 
474537   4593

stops=10
16337 sites
Maximum number of observations per site: 10 
Mean number of observations per site: 10 
Sites with at least one detection: 1385 

Tabulation of y observations:
     0      1 
159319   4051

```{r}
null <- occu(~1 ~1, data=umf)
null.r <- occu(~(1|ct.st.rt) ~1, data=umf) #+ (1|StateAbb)

summary(null.r)

#obsCov
p.stop <- occu(~stop + (1|ct.st.rt) ~1, data=umf)
p.stop.f <- occu(~factor(stop) + (1|ct.st.rt) ~1, data=umf)
p.stop2 <- occu(~scale(stop) + I(scale(stop)^2) + (1|ct.st.rt) ~1, data=umf)

modSel(fitList(null.r, p.stop, p.stop2))

#siteCov
p.stemp <- occu(~scale(StartTemp.F) + (1|ct.st.rt) ~1, data=umf)
# p.etemp <- occu(~EndTemp.F + (1|ct.st.rt) ~1, data=umf)

p.month <- occu(~Month + (1|ct.st.rt) ~1, data=umf)
p.jday <- occu(~scale(jday) + (1|ct.st.rt) ~1, data=umf)
p.jday2 <- occu(~scale(jday) + jday2 + (1|ct.st.rt) ~1, data=umf)

p.wind <- occu(~scale(StartWind) + (1|ct.st.rt) ~1, data=umf)
p.sky <- occu(~scale(StartSky) + (1|ct.st.rt) ~1, data=umf)

p.time <- occu(~scale(StartTime) + (1|ct.st.rt) ~1, data=umf)
# p.time2 <- occu(~StartTime + I(StartTime*StartTime) + (1|ct.st.rt) ~1, data=umf)
p.time2 <- occu(~scale(StartTime) + StartTime2 + (1|ct.st.rt) ~1, data=umf)

#model
p.fl <- fitList(null, null.r,
                p.stop, #per minute within a survey
                p.stemp, #p.etemp, 
                p.month, p.jday, p.jday2,
                p.wind, p.sky, #site covariates
                p.time, p.time2) #timing
modSel(p.fl)
#         nPars      AIC   delta    AICwt cumltvWt
# p.stop      3 18014.99    0.00  1.0e+00     1.00
# p.jday      3 18855.01  840.01 3.9e-183     1.00
# p.jday2     4 18856.43  841.43 1.9e-183     1.00
# p.stemp     3 18859.01  844.02 5.3e-184     1.00
# p.wind      3 18861.98  846.99 1.2e-184     1.00
# p.time2     4 18904.18  889.18 8.2e-194     1.00
# p.time      3 18904.79  889.79 6.1e-194     1.00
# p.sky       3 18965.18  950.18 4.7e-207     1.00

# null.r      2 18965.60  950.61 3.8e-207     1.00
# p.month     4 18965.97  950.98 3.1e-207     1.00
# null        2 26134.43 8119.44  0.0e+00     1.00

Sys.time()
p.global <- occu(~scale(stop) + scale(jday) + scale(StartTemp.F) + scale(StartWind) + #scale(StartSky) +
                   scale(StartTime) + StartTime2 + 
                   (1|ct.st.rt) ~1, data=umf)
Sys.time()
```

## coefficients and marginal plots for detection variables
```{r}
#these look kind of weird because need to set random effect variables
plotEffects(p.stop, "det", "stop")
plotEffects(p.stemp, "det", "StartTemp.F")
# plotEffects(p.etemp, "det", "EndTemp.F")
plotEffects(p.wind, "det", "StartWind")
plotEffects(p.sky, "det", "StartSky")
plotEffects(p.month, "det", "Month")
plotEffects(p.time, "det", "StartTime")
plotEffects(p.time2, "det", "StartTime")

p.coefs <- vals.coefs(p.global) #choose model to extract coefficients from

c.plot <- ggplot(data=p.coefs[!p.coefs$name %in% "Int",], aes(x=name, y=est, group=p, shape=p)) +
  # c.plot <- ggplot(data=coefs[coefs$name != "INT",], aes(x=name, y=est)) +
  geom_point(size=3, position=position_dodge(width=0.5)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.5)) +
  # geom_errorbar(aes(ymin=est-se, ymax=est+se), width=0, lwd=1, position=position_dodge(width=0.5)) +
  geom_hline(aes(yintercept=0), lty="dotted") +
  xlab("parameter") + ylab("estimate") +
  coord_flip() +
  theme_bw(base_size=16) + theme(panel.grid=element_blank()) + facet_wrap(.~p, scales="free", ncol=1)
c.plot

ggsave(c.plot, filename="../Figures/occupancy_BBS/CONI_stacked_occupancy_det_coefficients.tiff", height=6, width=6, units="in", dpi=400, compression="lzw")

# ###########
# #manually plot marginals in ggplot2
nd.mean <- site.stacked %>% mutate_if(is.numeric, mean)
nd.mean <- nd.mean[100,] #only take one row since all rows are the same (which row matters, since random effect by route and state)
nd.mean$stop <- 1

#
#plot these with p.global2, need to pull out scale() from within I() to plot these
plot.marginal(sp="CONI", p.global, "jday", "Julian day", type="det")
plot.marginal(sp="CONI", p.global, "StartTemp.F", "start temperature (F)", type="det")
plot.marginal(sp="CONI", p.global, "StartWind", "wind", type="det")

#Start time is a little weird since using chron package
Sys.setenv(TZ='GMT')

type <- "det"
st.time <- plot.marginal(sp="CONI", p.global, "StartTime", "Start Time", type="det")
  st.time + scale_x_chron(format="%H:%M", n=10)
  ggsave(filename=paste("../Figures/occupancy_BBS/", toupper(sp), "_stacked_occupancy_",type,"_", "StartTime",".tiff", sep=""), height=6, width=6, unit="in", dpi=300, compression="lzw")

#for stop
mod <- p.global; par <- "stop"; par.name <- "stop number"; type <- "det"
nd.og <- nd.mean[,names(nd.mean) != par] %>% slice(rep(1:n(), times=10))
nd <- data.frame(nd.og, par=1:10)
names(nd) <- gsub(names(nd), pattern="par", replacement=par)
marginal <- unmarked::predict(mod, type=type, newdata=nd)
o <- ggplot(marginal, aes(x=nd[,par], y=Predicted)) +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.5, fill="grey75") +
  ylab(paste("Predicted probability of", ylabel[ylabel$type == type,]$label)) +ylim(0,1) +
  xlab(par.name) + scale_x_continuous(breaks=1:10) +
  geom_line() + theme_bw(base_size=20) + theme(panel.grid=element_blank())
print(o)
ggsave(o, filename=paste("../Figures/occupancy_BBS/CONI_stacked_occupancy_",type,"_", par,".tiff", sep=""), height=6, width=6, unit="in", dpi=300, compression="lzw")


```


## run occupancy models for stacked single season occupancy models
```{r}

o.null <- occu(~1 ~(1|ct.st.rt), data=umf) #+ (1|StateAbb)

o.state <- occu(~1 ~ StateAbb + (1|ct.st.rt), data=umf)
o.bcr <- occu(~1 ~ BCR + (1|ct.st.rt), data=umf) 
o.year <- occu(~1 ~ Year + (1|ct.st.rt), data=umf)

o.lat <- occu(~1 ~ scale(Latitude) + (1|ct.st.rt), data=umf)
o.lat2 <- occu(~1 ~ scale(Latitude) + I(scale(Latitude)* scale(Latitude)) + (1|ct.st.rt), data=umf)

o.long <- occu(~1 ~ scale(Longitude) + (1|ct.st.rt), data=umf)
o.long2 <- occu(~1 ~ scale(Longitude) + I(scale(Longitude)* scale(Longitude)) + (1|ct.st.rt), data=umf)

o.fl <- fitList(o.lat, o.lat2, o.long, o.long2)
modSel(o.fl)

o.fl <- fitList(null, o.bcr, o.state, o.year,
                o.lat, o.lat2, o.long, o.long2)
modSel(o.fl)



```

## coefficients and marginal plots for occupancy variables
```{r}

```

