---
title: "2b_BBS_stacked_single-season occupancy_CWWI"
author: "Marie I. Tosa"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load packages

require(unmarked)
require(plyr)
require(dplyr)
require(tidyr)
require(chron) #for times function

require(ggplot2)
require(ggpubr)
```


## Create functions to plot figures
```{r}
vals.coefs <- function(mod)
{
coefs <- coef(mod) %>% tibble::enframe() #extract coefficients for covariates
se <- SE(mod) %>% tibble::enframe() #extract SEs for covariates

ci.psi <- confint(mod, type="state") %>% tibble::enframe()
ci.det <- confint(mod, type="det") %>% tibble::enframe()
ci <- rbind(ci.psi, ci.det) #combine psi and det CIs
ci <- data.frame(name=ci$name, lower=ci$value[,1], upper=ci$value[,2]) #need to split up ci so can read as data frame with lower and upper bounds

coefs <- merge(coefs, se, by="name") #merge coefficient values and ses
coefs <- merge(coefs, ci) #merge coefs with CIs
coefs <- data.frame(coefs)
names(coefs) <- c("par", "est", "se", "lower","upper")

coefs$p <- gsub(substr(coefs$par, 0,3), pattern="p\\([A-z]", replacement="det")
coefs$name <- gsub(coefs$par, pattern="scale\\(", replacement="")
coefs$name <- gsub(coefs$name, pattern="p\\(", replacement="")
coefs$name <- gsub(coefs$name, pattern="psi\\(", replacement="")
coefs$name <- gsub(coefs$name, pattern="\\)\\)", replacement="")
coefs$name <- gsub(coefs$name, pattern="Int\\)", replacement="Int")
coefs$name <- gsub(coefs$name, pattern="\\)", replacement="")
coefs$name <- gsub(coefs$name, pattern="I\\(", replacement="")
return(coefs)
}

ylabel <- data.frame(type=c("det","state"), label=c("detection","occupancy"))
plot.marginal <- function(sp, mod, par, par.name, type)
{
  nd.og <- nd.mean[,names(nd.mean) != par] %>% slice(rep(1:n(), times=1000))
  nd <- data.frame(nd.og,
                   par=seq(min(site.stacked[,par]), max(site.stacked[,par]), length.out=1000))
  
  names(nd) <- gsub(names(nd), pattern="par", replacement=par)
  marginal <- predict(mod, type=type, newdata=nd)
  
  nd$StartTime <- times(nd$StartTime) #convert start time back to character

  o <- ggplot(marginal, aes(x=nd[,par], y=Predicted)) +
    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.5, fill="grey75") +
    geom_rug(data=site.stacked, aes(x=site.stacked[,par], y=0)) +
    ylim(0,1) +
    xlab(par.name) + ylab(paste("Predicted probability of", ylabel[ylabel$type == type,]$label)) +
    geom_line() + theme_bw(base_size=20) + theme(panel.grid=element_blank())
  print(o)
  ggsave(o, filename=paste("../Figures/occupancy_BBS/", sp, "_stacked_occupancy_",type,"_", par,".tiff", sep=""), height=6, width=6, unit="in", dpi=300, compression="lzw")
  return(o)
}

```

## load data
Created detection history data files for each species in 2a_BBS_nightjar_detection_timing.Rmd

```{r}
sp <- "cwwi"

data <- read.table(paste("../Data_processed/BBS_2022/BBS_", sp, ".txt", sep=""), sep=",", header=T)
summary(data)

table(data$Year)
table(data$Month)

#fix start time
data[data$StartTime == 0,]$StartTime <- NA
data$StartTime <- times(paste(substr(data$StartTime, 0,1), ":", substr(data$StartTime, 2,3), ":00", sep=""))

data$StartTemp.F <- as.numeric(data$StartTemp.F)
data$EndTemp.F <- as.numeric(data$EndTemp.F)

#remove rows with incomplete data
data <- data[!is.na(data$StartTemp.F),]
# data <- data[!is.na(data$EndTemp.F),]

data <- data[!is.na(data$StartTime),]

#remove years before 1997 since hardly any detections then
data <- data[data$Year >= 1997,]


#####
data$StateAbb <- factor(data$StateAbb)
data$Month <- factor(data$Month)
# data$Year <- factor(data$Year)
data$BCR <- factor(data$BCR)

```

## create umf for stacked single-season occupancy model
```{r}
sp.stacked <- data[,grep(names(data), pattern="Stop[0-9]")]
sp.stacked[sp.stacked > 1 & !is.na(sp.stacked)] <- 1 #convert > 1 to 1 (presence/absence instead of counts)

#check detection histories
unique(sp.stacked)

colSums(sp.stacked) #can we cut off detections anywhere? aka reduce the number of stops
ggplot(data=data.frame(colSums(sp.stacked)), aes(x=1:50, y=colSums.sp.stacked.)) + geom_point(size=3) + 
  geom_hline(aes(yintercept=0), col="grey75", lty="dashed") +
  ylab("Num. detections") + xlab("stop") + theme_bw(base_size=20) + theme(panel.grid = element_blank())
ggsave(filename = "../Figures/occupancy_BBS/CWWI_detectionsbystop.tiff", height=3, width=9, units="in", dpi=400, compression="lzw")

sp.stacked <- sp.stacked[,1:8] #only take the first 10 stops since others are surveys mid-day

######
#siteCovs
site.stacked <- data[,!names(data) %in% grep(names(data), pattern="Stop", value=T)] #siteCovs

#add in transformations for variables, if any
#
#
site.stacked$StartTime2 <- scale(site.stacked$StartTime)^2
site.stacked$jday2 <- scale(site.stacked$jday)^2

site.stacked$Latitude2 <- scale(site.stacked$Latitude)^2
site.stacked$Longitude2 <- scale(site.stacked$Longitude)^2

######
#obsCovs
stop.df <- matrix(rep(1:length(names(sp.stacked)), times=nrow(sp.stacked)), byrow=T, ncol=length(names(sp.stacked))) #add obsCovs, dimensions the same as y data

umf <- unmarkedFrameOccu(y=sp.stacked, siteCovs=site.stacked, obsCovs=list(stop=stop.df))
head(umf)

# tiff(filename=paste("../Figures/occupancy_BBS/",sp,"dethist_BBS_stacked.tiff", sep=""), height=10, width=8, units="in", compression="lzw", res=400)
# plot(umf)
# dev.off()

plot(umf)
summary(umf)
```


## run detection models for stacked single season occupancy models
16337 sites
Maximum number of observations per site: 8 
Mean number of observations per site: 8 
Sites with at least one detection: 2751 

Tabulation of y observations:
     0      1 
124972   5724 


```{r}
null <- occu(~1 ~1, data=umf)
null.r <- occu(~(1|ct.st.rt) + (1|StateAbb) ~1, data=umf)

#obsCov
p.stop <- occu(~scale(stop) + (1|ct.st.rt) + (1|StateAbb) ~1, data=umf)

#siteCov
p.stemp <- occu(~scale(StartTemp.F) + (1|ct.st.rt) + (1|StateAbb) ~1, data=umf)
# p.etemp <- occu(~EndTemp.F + (1|ct.st.rt) + (1|StateAbb) ~1, data=umf)

p.month <- occu(~Month + (1|ct.st.rt) ~1, data=umf)
p.jday <- occu(~scale(jday) + (1|ct.st.rt) ~1, data=umf)
p.jday2 <- occu(~scale(jday) + jday2 + (1|ct.st.rt) ~1, data=umf)

p.wind <- occu(~scale(StartWind) + (1|ct.st.rt) + (1|StateAbb) ~1, data=umf)
p.sky <- occu(~scale(StartSky) + (1|ct.st.rt) + (1|StateAbb) ~1, data=umf)

p.time <- occu(~scale(StartTime) + (1|ct.st.rt) + (1|StateAbb) ~1, data=umf)
# p.time2 <- occu(~StartTime + I(StartTime*StartTime) + (1|ct.st.rt) + (1|StateAbb) ~1, data=umf)
p.time2 <- occu(~scale(StartTime) + StartTime2 + (1|ct.st.rt) + (1|StateAbb) ~1, data=umf)

#model
p.fl <- fitList(null, null.r,
                p.stop, #per minute within a survey
                p.stemp, #p.etemp, 
                p.month, p.jday, p.jday2,
                p.wind, p.sky, #site covariates
                p.time, p.time2) #timing
modSel(p.fl)
#         nPars      AIC    delta AICwt cumltvWt
# p.stop      3 22634.91     0.00  1.00     1.00
# p.time2     4 29531.19  6896.28  0.00     1.00
# p.time      3 29531.87  6896.97  0.00     1.00
# p.stemp     3 29783.87  7148.96  0.00     1.00

# null.r      2 29842.83  7207.92  0.00     1.00
# p.sky       3 29843.22  7208.31  0.00     1.00
# p.wind      3 29843.95  7209.05  0.00     1.00
# p.jday2     4 30286.13  7651.23  0.00     1.00
# p.jday      3 30290.25  7655.35  0.00     1.00
# p.month     4 30323.30  7688.40  0.00     1.00
# null        2 39426.30 16791.39  0.00     1.00

p.global <- occu(~scale(stop) + scale(StartTemp.F) + scale(StartTime) + StartTime2 +
                   (1|ct.st.rt) + (1|StateAbb) ~1, data=umf)
p.global1 <- occu(~scale(stop) + scale(StartTemp.F) + scale(StartTime) + I(scale(StartTime)^2) +
                   (1|ct.st.rt) + (1|StateAbb) ~1, data=umf)

```

## coefficients and marginal plots for detection variables
```{r}
#these look kind of weird because need to set random effect variables
plotEffects(p.stop, "det", "stop")
plotEffects(p.stemp, "det", "StartTemp.F")
plotEffects(p.etemp, "det", "EndTemp.F")
plotEffects(p.wind, "det", "StartWind")
plotEffects(p.sky, "det", "StartSky")
plotEffects(p.month, "det", "Month")
plotEffects(p.time, "det", "StartTime")
plotEffects(p.time2, "det", "StartTime")

p.coefs <- vals.coefs(p.global) #choose model to extract coefficients from

c.plot <- ggplot(data=p.coefs[!p.coefs$name %in% "Int",], aes(x=name, y=est, group=p, shape=p)) +
  # c.plot <- ggplot(data=coefs[coefs$name != "INT",], aes(x=name, y=est)) +
  geom_point(size=3, position=position_dodge(width=0.5)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.5)) +
  # geom_errorbar(aes(ymin=est-se, ymax=est+se), width=0, lwd=1, position=position_dodge(width=0.5)) +
  geom_hline(aes(yintercept=0), lty="dotted") +
  xlab("parameter") + ylab("estimate") +
  coord_flip() +
  theme_bw(base_size=16) + theme(panel.grid=element_blank()) + facet_wrap(.~p, scales="free", ncol=1)
c.plot

ggsave(c.plot, filename="../Figures/occupancy_BBS/CWWI_stacked_occupancy_det_coefficients.tiff", height=6, width=6, units="in", dpi=400, compression="lzw")

# ###########
# #manually plot marginals in ggplot2
nd.mean <- site.stacked %>% mutate_if(is.numeric, mean)
nd.mean <- nd.mean[100,] #only take one row since all rows are the same (which row matters, since random effect by route and state)
nd.mean$stop <- 1

#
#plot these with p.global2, need to pull out scale() from within I() to plot these
plot.marginal(sp="CWWI", p.global, "StartTemp.F", "start temperature (F)", type="det")
# plot.marginal(sp="CWWI", p.global, "StartWind", "wind", type="det")

#Start time is a little weird since using chron package
Sys.setenv(TZ='GMT')

type <- "det"
st.time <- plot.marginal(sp="CWWI", p.global1, "StartTime", "Start Time", type="det")
  st.time + scale_x_chron(format="%H:%M", n=10)
  ggsave(filename=paste("../Figures/occupancy_BBS/", toupper(sp), "_stacked_occupancy_",type,"_", "StartTime",".tiff", sep=""), height=6, width=6, unit="in", dpi=300, compression="lzw")

#for stop
mod <- p.global; par <- "stop"; par.name <- "stop number"; type <- "det"
nd.og <- nd.mean[,names(nd.mean) != par] %>% slice(rep(1:n(), times=10))
nd <- data.frame(nd.og, par=1:10)
names(nd) <- gsub(names(nd), pattern="par", replacement=par)
marginal <- unmarked::predict(mod, type=type, newdata=nd)
o <- ggplot(marginal, aes(x=nd[,par], y=Predicted)) +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.5, fill="grey75") +
  ylab(paste("Predicted probability of", ylabel[ylabel$type == type,]$label)) +ylim(0,1) +
  xlab(par.name) + scale_x_continuous(breaks=1:10) +
  geom_line() + theme_bw(base_size=20) + theme(panel.grid=element_blank())
print(o)
ggsave(o, filename=paste("../Figures/occupancy_BBS/CWWI_stacked_occupancy_",type,"_", par,".tiff", sep=""), height=6, width=6, unit="in", dpi=300, compression="lzw")


```


## run occupancy models for stacked single season occupancy models
```{r}
null <- occu(~1 ~1, data=umf)
null.r <- occu(~1 ~(1|ct.st.rt), data=umf)

o.state <- occu(~1 ~ StateAbb + (1|ct.st.rt), data=umf)
o.bcr <- occu(~1 ~ BCR + (1|ct.st.rt), data=umf)

o.lat <- occu(~1 ~ scale(Latitude) + (1|ct.st.rt), data=umf)
o.lat2 <- occu(~1 ~ scale(Latitude) + I(scale(Latitude)^2) + (1|ct.st.rt), data=umf)
o.long <- occu(~1 ~ scale(Longitude) + (1|ct.st.rt), data=umf)
o.long2 <- occu(~1 ~ scale(Longitude) + I(scale(Longitude)^2) + (1|ct.st.rt), data=umf)

o.year <- occu(~1 ~ scale(Year) + (1|ct.st.rt), data=umf)
o.year.f <- occu(~1 ~ factor(Year) + (1|ct.st.rt), data=umf)

o.fl <- fitList(null, null.r,
                o.state, o.bcr, o.lat, o.lat2, o.long, o.long2,
                o.year, o.year.f)
modSel(o.fl)

#          nPars      AIC   delta    AICwt cumltvWt
# o.lat2       4 30732.77    0.00  1.0e+00     1.00
# o.state     18 30764.69   31.92  1.2e-07     1.00 #insane SEs
# o.bcr        8 30828.90   96.13  1.3e-21     1.00
# o.lat        3 30829.07   96.30  1.2e-21     1.00
# o.long2      4 31079.22  346.45  5.9e-76     1.00
# o.long       3 31085.06  352.29  3.2e-77     1.00
# o.year.f    25 31096.39  363.62  1.1e-79     1.00
# o.year       3 31153.91  421.15  3.5e-92     1.00
# null.r       2 31214.84  482.07 2.1e-105     1.00
# null         2 39426.30 8693.53  0.0e+00     1.00

o.global <- occu(~1 ~ scale(Latitude) + Latitude2 + scale(Longitude) + Longitude2 +
                   scale(Year) + (1|ct.st.rt), data=umf)
o.global1 <- occu(~1 ~ scale(Latitude) + I(scale(Latitude)^2) + scale(Longitude) + Longitude2 +
                   scale(Year) + (1|ct.st.rt), data=umf)
o.global2 <- occu(~1 ~ scale(Latitude) + Latitude2 + scale(Longitude) + I(scale(Longitude)^2) +
                   scale(Year) + (1|ct.st.rt), data=umf)


```

## coefficients and marginal plots for occupancy variables
```{r}
nd.mean <- site.stacked %>% mutate_if(is.numeric, mean)
nd.mean <- nd.mean[200,] #only take one row since all rows have the same numeric values (but st_id matters)

# plot.marginal(sp="CWWI", o.year, par="Year", par.name="Year", type="state")
plot.marginal(sp="CWWI", o.global, par="Year", par.name="Year", type="state")

plot.marginal(sp="CWWI", o.global1, par="Latitude", par.name="Latitude", type="state")
plot.marginal(sp="CWWI", o.global2, par="Longitude", par.name="Longitude", type="state")


###########
#bcr
# mod <- o.global
mod <- o.bcr

require(sf)
BCR.ST <- read_sf(dsn="../../GIS", layer="BCR-state") #area in mi^2
BCR <- BCR.ST %>% group_by(BCR) %>% summarise(AreaSqMi = sum(AreaSqMi, na.rm=T), nstates=n())

nd.centroid <- st_centroid(BCR) %>% st_coordinates()
st_id <- unique(site.stacked[,c("BCR","ct.st.rt","StateAbb")]) %>%
  group_by(BCR) %>%
  arrange(ct.st.rt) %>%
  filter(row_number()==1)

nd <- cbind(BCR[,c("BCR")] %>% st_drop_geometry, nd.centroid)
names(nd) <- c("BCR","Longitude","Latitude")

nd <- merge(nd, st_id, by=c("BCR"), all.x=T)

nd$Latitude2 <- scale(nd$Latitude)^2
nd$Longitude2 <- scale(nd$Longitude)^2
nd$BCR <- factor(nd$BCR)
nd$Year <- factor(2018)

marginal <- predict(mod, type="state", newdata=nd)
m.data <- cbind(nd, marginal)
m.data$BCR <- factor(m.data$BCR, levels=rev(c(31,27,29,28,30,13,14)))

ggplot(m.data, aes(x=BCR, y=Predicted)) +
  geom_hline(aes(yintercept=0), lty="dashed", col="grey50") + geom_hline(aes(yintercept=1), lty="dashed", col="grey50") +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.25)) +
  geom_point(size=5, col="black", position=position_dodge(width=0.25)) +
  xlab("BCR") + ylab(paste("Predicted probability of occupancy")) + ylim(c(0,1)) +
  theme_bw(base_size=20) + theme(panel.grid=element_blank())

########
#bcr with year
u.year <- unique(site.stacked[,c("BCR","Year")])
nd <- cbind(BCR[,c("BCR")] %>% st_drop_geometry, nd.centroid)
names(nd) <- c("BCR","Longitude","Latitude")

nd <- merge(nd, st_id, by=c("BCR"), all.x=T)

nd$Latitude2 <- scale(nd$Latitude)^2
nd$Longitude2 <- scale(nd$Longitude)^2

nd$BCR <- factor(nd$BCR)

nd <- merge(nd, u.year, by="BCR", all.x=T)

marginal <- predict(mod, type="state", newdata=nd)
m.data <- cbind(nd, marginal)
# m.data <- m.data[order(m.data$state),]
# m.data$state <- factor(m.data$state, levels=rev(c("FL","GA","SC","NC","VA","WV","MD","DE","NJ","PA","NY","CT","MA","VT","NH","ME")))
m.data$BCR <- factor(m.data$BCR, levels=rev(c(31,27,29,28,30,13,14)))

ggplot(m.data, aes(x=Year, y=Predicted, col=BCR, fill=BCR, group=BCR)) +
  geom_hline(aes(yintercept=0), lty="dashed", col="grey50") + geom_hline(aes(yintercept=1), lty="dashed", col="grey50") +
  # geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.75)) +
  # geom_point(size=3, position=position_dodge(width=0.75)) +
  geom_ribbon(aes(ymin=lower, ymax=upper), col=NA, lwd=1, alpha=0.1) +
  geom_line() + facet_wrap(~BCR, ncol=2) +
  scale_color_brewer(palette="Set2") + scale_fill_brewer(palette="Set2") +
  xlab("Year") + ylab(paste("Predicted probability of occupancy")) + ylim(c(0,1)) +
  theme_bw(base_size=20) + theme(panel.grid=element_blank())
ggsave(filename="../Figures/occupancy_BBS/CWWI_stacked_occupancy_psi_BCR-year.tiff", height=6, width=16, units="in", dpi=400, compression="lzw")

range(m.data[m.data$BCR==31,]$Predicted)
range(m.data[m.data$BCR==27,]$Predicted)


```

