---
title: "3_NSNvsBBS"
author: "Marie I. Tosa"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#for GIS
require(sf)
require(raster)

#for plotting
require(ggplot2)
require(ggrepel)
require(viridis)
```

## Load data
```{r, echo=T}

AF_states <- st_read(dsn="../../GIS", layer="AF_states") 
# BCR <- read_sf(dsn="../../GIS", layer="BCR_AF_clip")
BCR.ST <- read_sf(dsn="../../GIS", layer="BCR-state") #area in mi^2

nsn.current <- read.csv("../Data_processed/summary_nsn_non-nsn_surveys_current_bcr-state.csv")
head(nsn.current)
bbsdensity <- read.csv("../Data_processed/sites_per_state-bcr_by_bbs_density.csv")
head(bbsdensity)
```
## randomly distribute points based on area of each BCR-state combo
```{r, echo=T}

totalarea <- sum(BCR.ST$AreaSqMi) #440286.7 mi^2
BCR.ST$prop_area <- BCR.ST$AreaSqMi/totalarea

BCR.ST$nsn1k.area <- round(BCR.ST$prop_area * 1000, digits=0)
BCR.ST$nsn2k.area <- round(BCR.ST$prop_area * 2000, digits=0)
BCR.ST$nsn3k.area <- round(BCR.ST$prop_area * 3000, digits=0)
```

bbsdensity$meanBCR.ST = mean detections at stops along BBS survey route
for nsn1k, scaled meanBCR.ST so sum(meanBCR.ST) = 1, then multiplied by sample of preference (1000, 2000, and 3000 points)

## compare current routes to ideal routes from bbs density of nightjars and by density
```{r, echo=T}
BCR.ST.df <- BCR.ST
st_geometry(BCR.ST.df) = NULL #convert back to data frame

comp <- merge(BCR.ST.df[,c("ST","BCR","BCRNAME","PROVINCE_S","AreaSqMi","prop_area","nsn1k.area","nsn2k.area","nsn3k.area")],
        nsn.current[,c("state","BCR","numcurrentroutes","numyears","meansurveysperyear")], 
              by.x=c("ST","BCR"), by.y=c("state","BCR"), all=T)

names(bbsdensity) <- gsub(names(bbsdensity), pattern="k$", replacement="k.bbsdens")
comp <- merge(comp, bbsdensity,
              by.x=c("ST","BCR"), by.y=c("StateAbb","BCR"), all=T)

# comp <- merge(nsn.current, bbsdensity,
#               by.x=c("state","BCRNumber"), by.y=c("StateAbb","BCR"), all=T)
# 
# #for adding BCRName if BCRName was NA and had a BCRNumber
# for(b in unique(comp[is.na(comp$BCRName) & !is.na(comp$BCRNumber),]$BCRNumber))
# {
#   print(b)
#   comp[is.na(comp$BCRName) & comp$BCRNumber %in% b,]$BCRName <- 
#     BCR[BCR$BCRNumber %in% b,]$BCRName
# }

comp[is.na(comp$BCRNAME),]$BCRNAME <- ""
comp[is.na(comp)] <- 0
comp[comp$BCR == 0,]$BCR <- NA

#convert number of surveys to number of routes
#just do with nsn.2k.area
comp$nsn2k.area.routes <- NA
comp[!is.na(comp$BCR),]$nsn2k.area.routes <- round(comp[!is.na(comp$BCR),]$nsn2k.area/10, digits=0)

comp$nsn2k.bbsdens.routes <- NA
comp[!is.na(comp$BCR),]$nsn2k.bbsdens.routes <- round(comp[!is.na(comp$BCR),]$nsn2k.bbsdens/10, digits=0)

comp$nsn3k.area.routes <- NA
comp[!is.na(comp$BCR),]$nsn3k.area.routes <- round(comp[!is.na(comp$BCR),]$nsn3k.area/10, digits=0)

write.csv(comp, file="../Data_processed/sampling_comparison_nsn_bbs.csv", row.names=F)
```

## Plot maps with currrent survey numbers and ideal
maps with current / weighted by bbs surveys / weighted by area of bcr-state
```{r, echo=T}

#plot
comp$plot_label <- paste(round(comp$meansurveysperyear, digits=0), comp$nsn2k.bbsdens, comp$nsn2k.area, sep="/")

comp$plot_label.routes <- paste(round(comp$numcurrentroutes, digits=0),
                                #pick one of these for 2k or 3k
                                # comp$nsn2k.bbsdens.routes, comp$nsn2k.area.routes, sep="/") 
                                comp$nsn3k.area.routes, sep="/")

comp.sf <- merge(BCR.ST[,c("BCR","ST","geometry")], comp, by=c("BCR","ST"), all=T)

comp.sf$ST <- factor(comp.sf$ST)
comp.sf$BCR <- factor(comp.sf$BCR)


comp.plot <- ggplot(data=comp.sf) +
  geom_sf(aes(fill=BCR), alpha=0.3) +
  scale_fill_brewer(palette="Set2") +
  scale_color_brewer(palette="Set2") +
  theme_bw(base_size=20) + xlab("") + ylab("") +
  theme(panel.grid=element_blank(), legend.position=c(0.85, 0.15))

comp.plot + 
  geom_text_repel(aes(label=plot_label, geometry=geometry, col=BCR), 
                  stat = "sf_coordinates")
ggsave(filename="../Figures/routes_per_bcr-state.tiff",
       height=15, width=15, units="in", dpi=400, compression="lzw")

comp.plot + 
  geom_sf_text(aes(label=plot_label.routes), col="black")
  # geom_text_repel(aes(label=plot_label.routes, geometry=geometry), col="black", 
  #                 stat = "sf_coordinates")
# ggsave(filename="../Figures/samples_per_bcr-state_2k.tiff", #pick one of these
ggsave(filename="../Figures/samples_per_bcr-state_3k.tiff",
       height=15, width=15, units="in", dpi=400, compression="lzw")

bbs.dens.plot <- ggplot(data=comp.sf) +
  geom_sf(aes(fill=meanBCR.ST), alpha=0.3) +
  scale_fill_viridis(option="turbo") +
  geom_text_repel(aes(label=round(meanBCR.ST, digits=2), geometry=geometry, col=meanBCR.ST), 
                  stat = "sf_coordinates") +
  scale_color_viridis(option="turbo") +
  theme_bw(base_size=20) + xlab("") + ylab("") +
  theme(panel.grid=element_blank(), legend.position=c(0.85, 0.15))
bbs.dens.plot
# ggsave(bbs.dens.plot, filename="../Figures/BBS/bbs_nightjar_density_bcr-state.tiff",
#        height=15, width=15, units="in", dpi=400, compression="lzw")

```