---
title: "4b nightjar survey tables by state"
author: "Marie Tosa, Anthony Roberts, Patrick Devers, US Fish and Wildlife Service"
date: "`r Sys.Date()`"
geometry: margin=1cm

output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning=F, message=F)

options(knitr.kable.NA = '')

require(knitr)

#for GIS
require(sf)
require(raster)
require(ggspatial)

#for plotting
require(ggplot2)
require(ggrepel)

AF_states <- read_sf(dsn="../../GIS", layer="AF_states")

BCR.ST <- read_sf(dsn="../../GIS", layer="BCR-state") #area in mi^2
unique(BCR.ST$ST)
BCR.ST$ST <- factor(BCR.ST$ST)
BCR.ST$BCR <- factor(BCR.ST$BCR)

s2 <- read.csv("../Data_processed/table_for_states_summary.csv")
s2[s2$selected == 1,]$selected <- "selected"
s2[s2$selected == 0,]$selected <- "alternate"
names(s2) <- c("ST","BCR","selected","num.exist","num.new","num.total")

s2 <- s2[order(s2$selected, decreasing=T),]
s2 <- s2[order(s2$BCR),]


data <- read.csv("../Data_processed/table_for_states.csv")
names(data) <- c("rand_rt_name","selected","existing_rt","Long.exist","Lat.exist","distance","X.rand","Y.rand","ST","BCR")
data$distance <- round(data$distance, digits=1)

#create function to print map for state
summary_table <- function(st)
{
  s2.st <- s2[s2$ST == st,]
  kable(s2.st, caption = "Number of random routes with existing routes vs number of random routes where new routes need to be established", row.names=F)
}

init_table <- function(st)
{
  #print(st)
  data.selected.st <- data[data$selected == 1 & data$ST == st,]
  kable(data.selected.st[,!names(data.selected.st) %in% c("selected","ST")], caption = "Initial nightjar survey routes", row.names=F)
}

alternate_table <- function(st)
{
  #print(st)
  data.alternate.st <- data[data$selected == 0 & data$ST == st,]
  kable(data.alternate.st[,!names(data.alternate.st) %in% c("selected","ST")], caption = "Alternate nightjar survey routes", row.names=F)
}

  
#base plot
plot <- ggplot() + 
  geom_sf(data=BCR.ST, aes(fill=BCR), alpha=0.5) +
  scale_fill_brewer(palette="Set2") +
  scale_shape_manual(values=c(21,24)) + scale_color_manual(values=c("black","grey50")) +
  xlab("") + ylab("") +
  theme_bw(base_size=20) + theme(panel.grid=element_blank())

init.map <- function(st)
{
  ex <- extent(AF_states[AF_states$ST == st,])

  data.st <- data[data$ST == st,]
  
  #create "alternate" column so easier to plot
  data.st$alternate <- 0
  data.st[data.st$selected == 0,]$alternate <- 1
  data.st$alternate <- factor(data.st$alternate)
  
  #create shape column, if there is a corresponding existing route
  data.st$shape <- "triangle"
  try(data.st[!is.na(data.st$Long.exist),]$shape <- "circle", silent=T)
  
  random.sf <- st_as_sf(data.st, coords=c("X.rand","Y.rand"), crs=4269)
  existing.sf <- st_as_sf(data.st[!is.na(data.st$Long.exist),], coords=c("Long.exist","Lat.exist"), crs=4269)

  #only include points that were selected
  plot + 
    geom_sf(data=random.sf[random.sf$selected == 1 & is.na(random.sf$Long.exist),], aes(col=alternate), pch=24, fill="black", size=3) + #triangle for random points without existing nearby
    geom_sf(data=random.sf[random.sf$selected == 1,], aes(col=alternate, shape=shape), fill="black", stroke=1.5) + 
    geom_sf(data=existing.sf[existing.sf$selected == 1,], col="steelblue", aes()) + 
    geom_text_repel(data=existing.sf[existing.sf$selected == 1,],
                    aes(label=existing_rt, geometry=geometry, col=alternate), stat = "sf_coordinates", col="steelblue", size=2) +
    coord_sf(xlim=c(ex@xmin, ex@xmax), ylim=c(ex@ymin-0.15, ex@ymax)) +
    annotation_scale(unit_category = "imperial") + #add scale bar
    theme(legend.position = "none")
}

all.map <- function(st)
{
  ex <- extent(AF_states[AF_states$ST == st,])
  data.st <- data[data$ST == st,]
  
  #create "alternate" column so easier to plot
  data.st$alternate <- 0
  data.st[data.st$selected == 0,]$alternate <- 1
  data.st$alternate <- factor(data.st$alternate)
  
  #create shape column, if there is a corresponding existing route
  data.st$shape <- "triangle"
  try(data.st[!is.na(data.st$Long.exist),]$shape <- "circle", silent=T)
  
  random.sf <- st_as_sf(data.st, coords=c("X.rand","Y.rand"), crs=4269)
  existing.sf <- st_as_sf(data.st[!is.na(data.st$Long.exist),], coords=c("Long.exist","Lat.exist"), crs=4269)

  #only include points that were selected
  plot + 
    geom_sf(data=random.sf[is.na(random.sf$Long.exist),], aes(col=alternate), pch=24, size=2) + #triangle for random points without existing nearby
    geom_sf(data=random.sf[random.sf$alternate==1,], aes(col=alternate, shape=shape), fill=NA, stroke=1.5) + 
    geom_sf(data=random.sf[random.sf$alternate==0,], aes(col=alternate, shape=shape), fill="black", stroke=1.5) + 
    
    geom_sf(data=existing.sf[existing.sf$alternate == 1,], col="steelblue", pch=1, stroke=1) + 
    geom_sf(data=existing.sf[existing.sf$alternate == 0,], col="steelblue", pch=16)  + 
    geom_text_repel(data=existing.sf,
                    aes(label=existing_rt, geometry=geometry, col=alternate), stat = "sf_coordinates", col="steelblue", size=2) +
    coord_sf(xlim=c(ex@xmin, ex@xmax), ylim=c(ex@ymin-0.15, ex@ymax)) +
    annotation_scale(unit_category = "imperial") +
    theme(legend.position = "none")
}

```
```{r, echo=F, results='asis', fig.dim=c(8.25,8.5)}
plot + theme(legend.position=c(0.85, 0.25))
```
\newpage

## Connectiut {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("CT")
init_table("CT")
alternate_table("CT")
init.map("CT")
all.map("CT")

```
\newpage

## Delaware {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("DE")
init_table("DE")
alternate_table("DE")
init.map("DE")
all.map("DE")
```
\newpage

## Florida {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("FL")
init_table("FL")
alternate_table("FL")
init.map("FL")
all.map("FL")
```
\newpage

## Georgia {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("GA")
init_table("GA")
alternate_table("GA")
init.map("GA")
all.map("GA")
```

\newpage

## Maine {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("ME")
init_table("ME")
alternate_table("ME")
init.map("ME")
all.map("ME")
```


\newpage

## Maryland {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("MD")
init_table("MD")
alternate_table("MD")
init.map("MD")
all.map("MD")
```

\newpage

## Massachusetts {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("MA")
init_table("MA")
alternate_table("MA")
init.map("MA")
all.map("MA")
```

\newpage

## New Hampshire {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("NH")
init_table("NH")
alternate_table("NH")
init.map("NH")
all.map("NH")
```

\newpage

## New Jersey {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("NJ")
init_table("NJ")
alternate_table("NJ")
init.map("NJ")
all.map("NJ")
```

\newpage

## New York {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("NY")
init_table("NY")
alternate_table("NY")
init.map("NY")
all.map("NY")
```

\newpage

## North Carolina {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("NC")
init_table("NC")
alternate_table("NC")
init.map("NC")
all.map("NC")
```

\newpage

## Pennsylvania {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("PA")
init_table("PA")
alternate_table("PA")
init.map("PA")
all.map("PA")
```

\newpage

## Rhode Island {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("RI")
init_table("RI")
alternate_table("RI")
init.map("RI")
all.map("RI")
```

\newpage

## South Carolina {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("SC")
init_table("SC")
alternate_table("SC")
init.map("SC")
all.map("SC")
```

\newpage

## New York {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("VT")
init_table("VT")
alternate_table("VT")
init.map("VT")
all.map("VT")
```

\newpage

## Virginia {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("VA")
init_table("VA")
alternate_table("VA")
init.map("VA")
all.map("VA")
```

\newpage

## West Virginia {.allowframebreaks}

```{r, echo=F, results='asis', fig.dim=c(8.25,8.25)}
summary_table("WV")
init_table("WV")
alternate_table("WV")
init.map("WV")
all.map("WV")
```
