---
title: "5_nsn_detectionhistories_2023-07-13"
author: "Marie I. Tosa"
date: "`r Sys.Date()`"
output: pdf_document
---

assumptions when combining data:
1. order of data entered for survey covariates was in order of route (e.g., stop 1 info entered before stop 2 info)
2. duplicate entries of observations per stop per species were assumed to be correct (e.g., if 2 observations entered for survey 1 stop 2, assumed multiple surveys were conducted at the stop). This is likely not true, but no way to correct this information without datasheets, etc.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)

#load packages
require(plyr)
require(dplyr)
require(tidyr)
require(lubridate)

require(sf)
require(raster)

require(ggplot2)

```

## load all data 
```{r}
AF_states <- st_read(dsn="../../GIS", layer="AF_states")

master.nsn <- read.table(file="../Data_processed/master_nsn_2023-07-24.txt", sep=",", header=T) #fixed multiple rows per survey issue, 8/17/2023 #latitude and longitude as flipped
master.nonnsn <- read.table(file="../Data_processed/master_nonnsn_2023-08-17.txt", sep=",", header=T) #added MA, 8/17/2023

nrow(master.nsn) #52,210
master.nsn <- master.nsn[!is.na(master.nsn$created_at),] #remove rows without observation data
nrow(master.nsn) #49,125

#convert t/f data to presence/absence data
master.nsn$stop_des <- ""
master.nsn$cars <- NA
master.nsn$dir <- NA

#convert date to julian day
master.nsn$start_date_j <- as.POSIXlt(master.nsn$start_date, tz="GMT", format="%Y-%m-%d")$yday

#convert time of day to something continuous
hours_in_day <- 24 #circumference = 2*pi
seconds_in_day <- 24*60*60 

master.nsn <- separate(master.nsn, col="start_hour", sep=":", into=c("hour","min","sec"), remove=F)

master.nsn$start_time_s <- sin((as.numeric(master.nsn$hour)*60*60 + 
                                  as.numeric(master.nsn$min)*60 + 
                                  as.numeric(master.nsn$sec))/seconds_in_day*(2*pi))
plot(as.numeric(master.nsn$hour), master.nsn$start_time_s) #looks good
#but some people did surveys in the middle of the day? From 11 am - 2 pm? 12 pm - 2 pm? 3pm?

master.nsn$start_time_c <- cos((as.numeric(master.nsn$hour)*60*60 + 
                                  as.numeric(master.nsn$min)*60 + 
                                  as.numeric(master.nsn$sec))/seconds_in_day*(2*pi))
plot(as.numeric(master.nsn$hour), master.nsn$start_time_c) #looks good

#give categorical value to hour
master.nsn$hour <- as.numeric(master.nsn$hour)
master.nsn$start_time_cat <- NA
master.nsn[master.nsn$hour < 4 & !is.na(master.nsn$hour),]$start_time_cat <- "night"
master.nsn[master.nsn$hour >= 4 & master.nsn$hour < 10 & !is.na(master.nsn$hour),]$start_time_cat <- "dawn"
master.nsn[master.nsn$hour >= 10 & master.nsn$hour < 17 & !is.na(master.nsn$hour),]$start_time_cat <- "day"
master.nsn[master.nsn$hour >= 17 & master.nsn$hour < 22 & !is.na(master.nsn$hour),]$start_time_cat <- "dusk"
master.nsn[master.nsn$hour >= 22 & !is.na(master.nsn$hour),]$start_time_cat <- "night"

nsn <- master.nsn[,c("route_id.char","stop_mit","state.rtid","name","start_date_j","start_month","hour","start_time_s","start_time_c","start_time_cat",
                     "Longitude","Latitude",
                     "start_year","stop_des","wind_id","sky_id","noise_id","cars","moon_visible","dir",
                     "code","min.1","min.2","min.3","min.4","min.5","min.6")]
names(nsn) <- c("id","Route.Stop","state","name","start_date_j","start_month","hour","start_time_s","start_time_c","start_time_cat",
                "Route.Latitude","Route.Longitude",
                "year","stop_des","wind","sky","noise","cars","moon","dir",
                "species","min.1","min.2","min.3","min.4","min.5","min.6")

#replace all t values with 1, all f values with 0
# nsn %>% mutate(min.1 = replace(min.1, min.1=="t", "1")) #only at one column
nsn$moon <- gsub(nsn$moon, pattern="\\N", replacement=NA)
nsn$sky <- gsub(nsn$sky, pattern="\\N", replacement=NA)
nsn$wind <- gsub(nsn$moon, pattern="\\N", replacement=NA)
nsn$noise <- gsub(nsn$noise, pattern="\\N", replacement=NA)

#already did this to min.1-min.6 in 0_Exploratory_analysis
nsn <- nsn %>% mutate_at(vars(wind,moon,min.1,min.2,min.3,min.4,min.5,min.6), function(x) replace(x, x=="t",1)) #all min columns at once
nsn <- nsn %>% mutate_at(vars(wind,moon,min.1,min.2,min.3,min.4,min.5,min.6), function(x) replace(x, x=="f",0)) #all min columns at once
nsn <- nsn %>% mutate_at(vars(wind,sky,min.1,min.2,min.3,min.4,min.5,min.6), function(x) as.integer(x)) #convert from character to integer
#OK with warning (converting \\N to integer -> NA)

#remove rows without lat/long information? Maybe don't need to do this
# nsn <- nsn[!is.na(nsn$Route.Latitude),] #23,317 rows with location information

nsn <- nsn[!is.na(nsn$year),] #remove rows without year info, no state info either
nrow(nsn) #46,901 rows with year information

#only AF data
nsn.af <- nsn[nsn$state %in% AF_states$ST,] #subset data for AF states
nrow(nsn.af) #18,814 rows with year information (didn't remove sites if they didn't have lat/long)

#  [1] "id"              "Route.Stop"      "state"           "name"            "created_at"      "updated_at"      "Route.Latitude"  "Route.Longitude"
#  [9] "year"            "stop_des"        "wind"            "sky"             "noise"           "cars"            "moon"            "dir"            
# [17] "species"         "min.1"           "min.2"           "min.3"           "min.4"           "min.5"           "min.6"           "TotalDetections"
# [25] "MaxDetections"

nsn.af$id <- as.character(nsn.af$id)
nsn.af$year <- as.integer(nsn.af$year) #OK to ignore warning about NAs
nsn.af$TotalDetections <- rowSums(nsn.af[,grep(names(nsn.af), pattern="min.")]) #min.1 through min.6, columns 19:24
suppressWarnings(nsn.af$MaxDetections <- apply(nsn.af[,grep(names(nsn.af), pattern="min.")], 1, max, na.rm=T)) #min.1 through min.6
# nsn.af[is.na(nsn.af$species),]$species <- "NONE"
nsn.af$noise <- as.integer(nsn.af$noise)

##########
#adjust non.nsn data
##########
master.nonnsn[master.nonnsn$start_hour %in% "",]$start_hour <- NA
master.nonnsn <- separate(master.nonnsn, col="start_hour", sep=":", into=c("hour","min","sec"), remove=F)
# master.nonnsn[!is.na(master.nonnsn$start_hour),]$sec <- 0

master.nonnsn$start_time_s <- sin((as.numeric(master.nonnsn$hour)*60*60 +
                                     as.numeric(master.nonnsn$min)*60 #+ as.numeric(master.nsn$sec)
                                     )/seconds_in_day*(2*pi))
plot(as.numeric(master.nonnsn$hour), master.nonnsn$start_time_s)

master.nonnsn$start_time_c <- cos((as.numeric(master.nonnsn$hour)*60*60 + 
                                  as.numeric(master.nonnsn$min)*60)/seconds_in_day*(2*pi))
plot(as.numeric(master.nonnsn$hour), master.nonnsn$start_time_c) #looks good

master.nonnsn$start_date_j <- as.POSIXlt(master.nonnsn$start_date, tz="GMT", format="%m/%d/%Y")$yday

master.nonnsn$hour <- as.numeric(master.nonnsn$hour)
table(master.nonnsn$hour)

master.nonnsn$start_time_cat <- NA
master.nonnsn[master.nonnsn$hour < 4 & !is.na(master.nonnsn$hour),]$start_time_cat <- "night"
master.nonnsn[master.nonnsn$hour >= 4 & master.nonnsn$hour < 10 & !is.na(master.nonnsn$hour),]$start_time_cat <- "dawn"
# master.nonnsn[master.nonnsn$hour >= 10 & master.nonnsn$hour < 17 & !is.na(master.nonnsn$hour),]$start_time_cat <- "day" #no times during the day
master.nonnsn[master.nonnsn$hour >= 17 & master.nonnsn$hour < 22 & !is.na(master.nonnsn$hour),]$start_time_cat <- "dusk"
master.nonnsn[master.nonnsn$hour >= 22 & !is.na(master.nonnsn$hour),]$start_time_cat <- "night"

non.nsn <- master.nonnsn[!is.na(master.nonnsn$year),] #14,138 rows
non.nsn <- non.nsn[,names(non.nsn) %in% names(nsn)]

############
#combine nsn and non.nsn survey information
############
nightjars <- bind_rows(nsn.af, non.nsn) #32,952 rows
nrow(nightjars)
nightjars$st_id <- paste(nightjars$state, nightjars$id, sep="--")

table(nightjars[is.na(nightjars$species),c("state","year")]) #see which surveys don't have species information

#delaware stop 8 was not surveyed. needs to stay NA. ME data not provided yet as of 8/1/2023. remove from data frame
nightjars <- nightjars[!is.na(nightjars$species),]

unique(nightjars$species)

#########
#remove duplicated survey entries
#don't need to do this any more. after changing the order of combining data tables?
# nightjars <- nightjars[order(nightjars$updated_at, nightjars$st_id),] #start with 27,599 rows
# nightjars <- nightjars[!duplicated(nightjars[,c("id","Route.Stop","state","name","Route.Latitude","Route.Longitude",
#                                                 "year","wind","sky","noise","species","min.1","min.2","min.3","min.4","min.5","min.6",
#                                                 "TotalDetections","MaxDetections","st_id")]),] #end with 21,242 rows

nightjars <- nightjars[order(nightjars$year, nightjars$st_id, nightjars$Route.Stop),] #22,339 rows

nightjars$Route.Latitude <- round(nightjars$Route.Latitude, digits=5)
nightjars$Route.Longitude <- round(nightjars$Route.Longitude, digits=5)
nrow(nightjars)

table(nightjars$state)
 #  CT   DE   FL   GA   MA   MD   ME   NC   NH   NY   PA   SC   VA   WV 
 # 658    9 2596  820 2776  231 6134 2492 1896 1886  587  886 1337   31

table(nightjars$species)
 # ANNI CONI CWWI EWPW MWPW NONE 
 #    2 1079 4639 8527    2 8090 

#make sure species = none gets detections of 0
#mostly ME sites
nightjars[nightjars$species %in% "NONE",]$min.1 <- 0
nightjars[nightjars$species %in% "NONE",]$min.2 <- 0
nightjars[nightjars$species %in% "NONE",]$min.3 <- 0
nightjars[nightjars$species %in% "NONE",]$min.4 <- 0
nightjars[nightjars$species %in% "NONE",]$min.5 <- 0
nightjars[nightjars$species %in% "NONE",]$min.6 <- 0

# table(nightjars$species)

#need to fix NAs in TotalDetections and MaxDetections for CT
nightjars$TotalDetections <- rowSums(nightjars[,grep(names(nightjars), pattern="min.", value=T)], na.rm = T)
nightjars$MaxDetections <- suppressWarnings(apply(nightjars[,grep(names(nightjars), pattern="min.", value=T)], 1, max, na.rm = T))

nightjars[nightjars$MaxDetections == 0,]$species <- "NONE"

#save file with all nsn and non-nsn data
write.table(nightjars, file="../Data_processed/master_alldata_2023-08-18.txt", sep=",") #22,339 observations

#IGNORE THIS. DON'T NEED TO DO THIS
###########
# site.info <- unique(nightjars[,c("Route.Latitude","Route.Longitude","year","st_id","id","Route.Stop","state","wind","sky","noise","cars","moon")])
# #create obs covariates for each minute?
# #repeat each row 6 times for each minute
# site.minute.info <- as.data.frame(lapply(site.info, rep, each=6))
# site.minute.info$minute <- rep(1:6, nrow(site.info))
# site.minute.info$site <- paste(site.minute.info$st_id, site.minute.info$Route.Stop, sep=".")

# #data frame of the unique st_id and Route.Stop per year
# #to determine which routes were surveyed each year (help fill in 0 detections of each species?)
# site.years <- unique(nightjars[,c("Route.Latitude","Route.Longitude","year","st_id","id","Route.Stop","state")])
# site.years$site <- paste(site.years$st_id, site.years$Route.Stop, sep=".")
# 
# #data frame of unique sites 
# sites <- unique(nightjars[,c("Route.Latitude","Route.Longitude","st_id","id","Route.Stop","state")]) #1842
# sites <- sites[!is.na(sites$Route.Latitude),]
# sites <- sites[!duplicated(sites[,c("st_id","Route.Stop")]),] #1760 routes and stops
# 
# nightjars[is.na(nightjars)] <- 0
# 
# st.bcr <- read.csv("../Data_processed/route_bcr_state.csv")
# 
# sites <- merge(sites, st.bcr, by.x=c("state","id"), by.y=c("state","route_id.char"), all.x=T) #add BCR number to sites
# site.years <- merge(site.years, st.bcr, by.x=c("state","id"), by.y=c("state","route_id.char"), all.x=T)
# 
# #save files
# write.table(nightjars, file="../Data_processed/occupancy_master_nightjars_2023-07-13.txt", sep=",", row.names=F)
# write.table(sites, file="../Data_processed/occupancy_sites_2023-07-13.txt", sep=",", row.names=F)
###########

```

## see how many surveys per stop per year
helps find duplicate survey entries
skip this
```{r}
# # table(nightjars[,c("id","Route.Stop","year")])
# route.stop.year.surveys <- ddply(nightjars, .(st_id, Route.Stop, year, species), summarize, numsurveys=length(TotalDetections))
# 
# route.stop.year.surveys <- route.stop.year.surveys[!is.na(route.stop.year.surveys$st_id),]
# 
# wide <- array(data=0, dim=c(12, length(unique(route.stop.year.surveys$st_id)), length(2007:2023))) # 12 Stops, 435 routes, 17 years
# 
# for(s in 1:12) #for each stop
# {
#   print(s)
#   s.route.year.survey <- route.stop.year.surveys[route.stop.year.surveys$Route.Stop == s & route.stop.year.surveys$species == "EWPW",]
#   s.route.year.survey <- s.route.year.survey[!is.na(s.route.year.survey$year),]
#   s.route.year.survey <- s.route.year.survey[order(s.route.year.survey$year),] #rearrange rows so columns are in order of year starting with 2007
#   s.route.year.survey <- rbind(data.frame(st_id="test", Route.Stop=13, species="CONI", year=2007:2023, numsurveys=0), s.route.year.survey)
#   s.wide <- pivot_wider(s.route.year.survey, values_from=numsurveys, names_from=year)
#   s.wide <- merge(s.wide, data.frame(routes=unique(route.stop.year.surveys$st_id)), by.x="st_id",by.y="routes", all.y=T) #make sure all sites are present
#   s.wide[is.na(s.wide)] <- 0
#   # print(dim(s.wide))
#   wide[s,,] <- as.matrix(s.wide[,4:20])
# }
# 
# # wide[1,,] #stop 1
# # wide[2,,] #stop 2
# 
# stop1 <- data.frame(wide[1,,])
# row.names(stop1) <- unique(route.stop.year.surveys$st_id)
# colnames(stop1) <- paste("y",2007:2023, sep="")
# 
# # nightjars[nightjars$st_id == "FL-250132" & nightjars$year == 2014 & nightjars$species == "CONI",]

```

## create detection history template
based on Cara's det_hist_prep.R code and Marie's 3_occupancy_detectionhistory_template_2022-10-05.R
skip this too since want stacked data format
```{r, echo=T}
# beg <- 2007
# end <- 2022
# 
# nMins <- 6
# 
# nC <- length(beg:end)*nMins
# 
# t <- paste(rep(paste("y", 2007:2022, sep=""), each=nMins), 
#            rep(paste("m", 1:6, sep=""), length(2007:2022)), sep="")
# 
# #make empty detection history template
# det.hist <- data.frame('site' = rep(paste(sites$st_id, sites$Route.Stop, sep="."), each=nC),
#                        'year-min'=rep(t, nrow(sites)))
# 
# head(det.hist)
# dim(det.hist) #one row for each site per year per minute
# nrow(sites) * nC
# 
# #need to turn detections of each species into long form
# nightjars$site <- paste(nightjars$st_id, nightjars$Route.Stop, sep="..")
# sites$site <- paste(sites$st_id, sites$Route.Stop, sep=".")
# 
# # site.years.min <- data.frame(s=rep(paste(site.years$site, site.years$year, sep="_"), each=6), minute=rep(1:6, nrow(site.years)))
# # site.years.min <- separate(site.years.min, col="s", sep="_", into=c("site","year"))
# 
# ewpw <- nightjars[nightjars$species == "EWPW",]
# # ewpw <- ewpw[,c("site","year","wind","sky","noise","cars","moon","min.1","min.2","min.3","min.4","min.5","min.6")]
# ewpw <- ewpw[,c("site","year","min.1","min.2","min.3","min.4","min.5","min.6")] #could use this for stacked occupancy models, but need to add in 0s of sites surveyed, but not detected
# 
# #for dynamic occupancy
# ewpw.long <- pivot_longer(data=ewpw, cols=min.1:min.6, names_to="minute", values_to="detection")
# ewpw.long$minute <- as.numeric(gsub(ewpw.long$minute, pattern="min.", replacement=""))
# 
# cwwi <- nightjars[nightjars$species == "CWWI",]
# coni <- nightjars[nightjars$species == "CONI",]
# 
# #merge with sites
# # ewpw.long.sites <- merge(ewpw.long, site.minute.info[,c("site","year","minute")], by=c("site","year","minute"), all.y=T) #necessary to add in 0s vs. NAs
# ewpw.long.sites <- merge(ewpw.long, site.info[,c("site","year")], by=c("site","year"), all.y=T) #necessary to add in 0s vs. NAs
# 
# ewpw.long.sites[is.na(ewpw.long.sites$detection),]$detection <- 0
# ewpw.long.sites$year.min <- paste("y", ewpw.long.sites$year, "m", ewpw.long.sites$minute, sep="")
# 
# #merge with det.hist
# ewpw.dethist.long <- merge(ewpw.long.sites[!duplicated(ewpw.long.sites[,c("site","year.min")]),], 
#                            #####take first non-duplicated row...come back to this#######
#                            det.hist, by=c("site","year.min"), all.y=T)
# 
# #convert from long to wide
# ewpw.dethist.wide <- pivot_wider(ewpw.dethist.long[,c("site","year.min","detection")], values_from=detection, names_from=year.min)
# 
# write.csv(ewpw.dethist.wide, file="../Data_processed/occupancy_detectionhistory_EWPW.csv", row.names=F)

```


## For stacked single seasson occupancy model detection histories
```{r}
# site.years <- unique(nightjars[,c("Route.Latitude","Route.Longitude","year","st_id","id","Route.Stop","state")])

st.bcr <- read.csv("../Data_processed/route_bcr_state.csv")

#check all state-id combos are in st.bcr
nightjars[!paste(nightjars$state, nightjars$id, sep="__") %in% paste(st.bcr$state, st.bcr$route_id.char, sep="__"), ]
unique(nightjars[!paste(nightjars$state, nightjars$id, sep="__") %in% paste(st.bcr$state, st.bcr$route_id.char, sep="__"), ]$st_id) #156 routes
# routes without location information (all from NSN data)
#   [1] "FL--256606"                  "FL--259296"                  "GA--279012"                  "NC--639319"                 
#   [5] "NC--639320"                  "NC--639321"                  "NC--639322"                  "NC--639323"                 
#   [9] "NC--639324"                  "NC--639325"                  "NC--639326"                  "NC--639327"                 
#  [13] "NC--639328"                  "NC--639329"                  "NC--639356"                  "NC--639358"                 
#  [17] "VA--881675"                  "VA--883203"                  "FL--256180"                  "FL--257284"                 
#  [21] "NC--639330"                  "NC--639331"                  "NC--639332"                  "NC--639333"                 
#  [25] "NC--639334"                  "NC--639335"                  "NC--639336"                  "NC--639337"                 
#  [29] "NC--639338"                  "NC--639339"                  "NC--639347"                  "NC--639348"                 
#  [33] "NC--639349"                  "VA--886865"                  "FL--256925"                  "MD--981235"                 
#  [37] "MD--981236"                  "NC--639340"                  "SC--801850"                  "VA--889503"                 
#  [41] "FL--251227"                  "FL--259009"                  "FL--259010"                  "FL--259011"                 
#  [45] "GA--279332"                  "NC--631636"                  "NC--638167"                  "NC--639019"                 
#  [49] "NC--639341"                  "NC--639342"                  "NC--639343"                  "PA--720071"                 
#  [53] "VA--889026"                  "FL--255710"                  "NC--639346"                  "NC--639354"                 
#  [57] "NC--639355"                  "VA--889506"                  "VA--889527"                  "FL--259354"                 
#  [61] "FL--259355"                  "FL--259358"                  "FL--259359"                  "FL--259363"                 
#  [65] "MA--Castle Neck, Ipswich"    "ME--5"                       "ME--6"                       "NC--639344"                 
#  [69] "PA--729722"                  "PA--729723"                  "SC--803736"                  "VA--889505"                 
#  [73] "VA--889509"                  "VA--889517"                  "VA--889518"                  "NC--639359"                 
#  [77] "SC--803738"                  "VA--889507"                  "VA--889520"                  "FL--259374"                 
#  [81] "NC--639370"                  "VA--889530"                  "GA--279336"                  "NC--639374"                 
#  [85] "SC--803739"                  "SC--803740"                  "VA--889528"                  "WV--909252"                 
#  [89] "GA--279339"                  "GA--279340"                  "GA--279342"                  "GA--279345"                 
#  [93] "MD--981237"                  "NC--639345"                  "NC--639377"                  "VA--889539"                 
#  [97] "GA--279348"                  "NC--639384"                  "PA--729730"                  "PA--729732"                 
# [101] "FL--259380"                  "GA--279349"                  "NC--639387"                  "NC--639389"                 
# [105] "PA--729735"                  "PA--729736"                  "PA--729737"                  "SC--803743"                 
# [109] "VA--889542"                  "VA--889549"                  "VA--889562"                  "GA--279350"                 
# [113] "NH--18"                      "PA--729739"                  "PA--729747"                  "PA--729748"                 
# [117] "SC--803747"                  "SC--803748"                  "VA--889560"                  "VA--889566"                 
# [121] "VA--889570"                  "VA--889571"                  "VA--889573"                  "VA--889574"                 
# [125] "VA--889575"                  "WV--909255"                  "FL--259386"                  "MA--Quabbin Islands by Boat"
# [129] "NC--639398"                  "NC--639399"                  "NC--639400"                  "NC--639401"                 
# [133] "NC--639402"                  "NC--639404"                  "PA--729750"                  "SC--803752"                 
# [137] "SC--803755"                  "VA--889581"                  "VA--889582"                  "VA--889586"                 
# [141] "FL--259384"                  "FL--259390"                  "FL--259393"                  "GA--279355"                 
# [145] "SC--803751"                  "SC--803758"                  "SC--803759"                  "SC--803760"                 
# [149] "SC--803761"                  "SC--803762"                  "SC--803763"                  "SC--803764"                 
# [153] "SC--803765"                  "VA--889551"                  "VA--889592"                  "WV--909256"   

nightjars <- merge(nightjars, st.bcr[,c("route_id.char","state","BCR")], by.x=c("state","id"), by.y=c("state","route_id.char"), all.x=T)
unique(nightjars[is.na(nightjars$BCR),]$st_id)  #check to see if all surveys were assigned a bcr, no :( these are sites without lat/long info

nightjars$site <- paste(nightjars$st_id, nightjars$Route.Stop, sep="..") #add stop to site name (st_id + Route.Stop)

#####
write.table(nightjars, file="../Data_processed/master_alldata_2023-08-18.txt", sep=",", row.names=F) #data with BCRs added
#id is the same as route

#####
#this isn't really useful for anything since can't merge it with stuff later
# #all sites that were surveyed each year
# site.years <- unique(nightjars[,c("site","Route.Latitude","Route.Longitude","year","st_id","id","Route.Stop","state","BCR",
#                                   "wind","sky","noise","cars","moon",
#                                   "start_date_j","start_month","start_time_s")])
# nrow(site.years) #27,927 site-years
# 
# #necessary to remove duplicated site-years
# covars <- site.years[,c("site","year","BCR","wind","sky","noise","cars","moon",
#                         "start_date_j","start_month","start_hour_s")] #site has route.stop embedded in it
# covars[duplicated(covars[,c("site","year")]),]


############
#remove 2nd and 3rd surveys from Maine data so it's comparable with other data
me.nightjars <- nightjars[nightjars$state == "ME",] #6136 rows

me.nightjars <- me.nightjars %>% #take first row of each group
  group_by(site, year, species) %>%
  arrange(start_date_j) %>%
  filter(row_number()==1)
nrow(me.nightjars) #3082 rows

nightjars <- nightjars[!nightjars$state %in% "ME",]
nightjars <- rbind(nightjars, me.nightjars) #add ME back in
nrow(nightjars) #19,287 rows

############
############
#separate tables for each species
#have to keep all covariates together with the data since there are duplicates of site observations (can't merge them together)
############
#EWPW
############
# ewpw <- nightjars[nightjars$species == "EWPW",]
# ewpw <- ewpw[,c("site","year","min.1","min.2","min.3","min.4","min.5","min.6")] #add 0 sites and siteCovs after

ewpw <- nightjars

ewpw1 <- ewpw %>% filter(species %in% c("EWPW")) #6908 rows #all ewpw rows
nrow(ewpw1)

# ewpw2 <- ewpw %>% filter(!species %in% c("EWPW")) %>% mutate_at(vars(matches("min")), ~0) #12379 rows #all rows with other species
ewpw2 <- ewpw %>% filter(!species %in% c("EWPW")) %>% mutate_at(vars(matches("min")), ~ifelse(is.na(.), NA, 0)) #if value is not NA, replace with 0
nrow(ewpw2)
ewpw2$species <- "EWPW"
ewpw2 <- ewpw2[!duplicated(ewpw2),] #remove duplicates

ewpw <- rbind(ewpw1, ewpw2)
ewpw <- ewpw[!duplicated(ewpw[,!names(ewpw) %in% "species"]),] #18,247 rows
nrow(ewpw)

# ewpw$species <- "EWPW"
suppressWarnings(ewpw$MaxDetections <- apply(ewpw[,grep(names(ewpw), pattern="min.")], 1, function(x) max(x, na.rm=T)))
suppressWarnings(ewpw$TotalDetections <- apply(ewpw[,grep(names(ewpw), pattern="min.")], 1, function(x) sum(x, na.rm=T)))

write.csv(ewpw, file="../Data_processed/occupancy_detectionhistory_stacked_EWPW.csv", row.names=F)

#########
#CWWI
#########
cwwi <- nightjars

cwwi1 <- cwwi %>% filter(species %in% c("CWWI")) #4630 rows #all cwwi rows
nrow(cwwi1)

# cwwi2 <- cwwi %>% filter(!species %in% c("CWWI")) %>% mutate_at(vars(matches("min")), ~0) #11,223 rows #all rows with other species
cwwi2 <- cwwi %>% filter(!species %in% c("CWWI")) %>% mutate_at(vars(matches("min")), ~ifelse(is.na(.), NA, 0))
cwwi2 <- cwwi2[!cwwi2$state %in% c("CT","MA"),] #need to remove CT and MA routes from CWWI since not surveyed
nrow(cwwi2)
cwwi2$species <- "CWWI"
cwwi2 <- cwwi2[!duplicated(cwwi2),] #remove duplicates

cwwi <- rbind(cwwi1, cwwi2)
cwwi <- cwwi[!duplicated(cwwi[,!names(cwwi) %in% "species"]),] #15,444 rows
nrow(cwwi)

# cwwi$species <- "CWWI"
suppressWarnings(cwwi$MaxDetections <- apply(cwwi[,grep(names(cwwi), pattern="min.")], 1, function(x) max(x, na.rm=T)))
suppressWarnings(cwwi$TotalDetections <- apply(cwwi[,grep(names(cwwi), pattern="min.")], 1, function(x) sum(x, na.rm=T)))

write.csv(cwwi, file="../Data_processed/occupancy_detectionhistory_stacked_CWWI.csv", row.names=F)


#########
#CONI
#########
coni <- nightjars

coni1 <- coni %>% filter(species %in% c("CONI")) #1,015 rows #all coni rows
nrow(coni1)

# coni2 <- coni %>% filter(!species %in% c("CONI")) %>% mutate_at(vars(matches("min")), ~0) #18,272 rows #all rows with other species
coni2 <- coni %>% filter(!species %in% c("CONI")) %>% mutate_at(vars(matches("min")), ~ifelse(is.na(.), NA, 0)) #replace values of all other species with 0 if not NA
nrow(coni2)
coni2$species <- "CONI"
coni2 <- coni2[!coni2$state %in% c("CT","MA"),] #need to remove CT and MA sites since only surveyed for EWPW
coni2 <- coni2[!duplicated(coni2),] #remove duplicates

coni <- rbind(coni1, coni2)
coni <- coni[!duplicated(coni[,!names(coni) %in% "species"]),] #15,378 rows
nrow(coni)

# coni$species <- "CONI"
suppressWarnings(coni$MaxDetections <- apply(coni[,grep(names(coni), pattern="min.")], 1, function(x) max(x, na.rm=T)))
suppressWarnings(coni$TotalDetections <- apply(coni[,grep(names(coni), pattern="min.")], 1, function(x) sum(x, na.rm=T)))

write.csv(coni, file="../Data_processed/occupancy_detectionhistory_stacked_CONI.csv", row.names=F)
```
