---
title: "6_nightjar surveys_single-season occupancy_stacked years_CWWI"
author: "Marie I. Tosa"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(unmarked)
require(plyr)
require(dplyr)
require(tidyr)

require(ggplot2)
require(ggpubr)
```

## Create functions to plot figures
```{r}
vals.coefs <- function(mod)
{
coefs <- coef(mod) %>% tibble::enframe() #extract coefficients for covariates
se <- SE(mod) %>% tibble::enframe() #extract SEs for covariates

ci.psi <- confint(mod, type="state") %>% tibble::enframe()
ci.det <- confint(mod, type="det") %>% tibble::enframe()
ci <- rbind(ci.psi, ci.det) #combine psi and det CIs
ci <- data.frame(name=ci$name, lower=ci$value[,1], upper=ci$value[,2]) #need to split up ci so can read as data frame with lower and upper bounds

coefs <- merge(coefs, se, by="name") #merge coefficient values and ses
coefs <- merge(coefs, ci) #merge coefs with CIs
coefs <- data.frame(coefs)
names(coefs) <- c("par", "est", "se", "lower","upper")

coefs$p <- gsub(substr(coefs$par, 0,3), pattern="p\\([A-z]", replacement="det")
coefs$name <- gsub(coefs$par, pattern="scale\\(", replacement="")
coefs$name <- gsub(coefs$name, pattern="p\\(", replacement="")
coefs$name <- gsub(coefs$name, pattern="psi\\(", replacement="")
coefs$name <- gsub(coefs$name, pattern="\\)\\)", replacement="")
coefs$name <- gsub(coefs$name, pattern="Int\\)", replacement="Int")
coefs$name <- gsub(coefs$name, pattern="\\)", replacement="")
coefs$name <- gsub(coefs$name, pattern="I\\(", replacement="")
return(coefs)
}

plot.marginal <- function(sp, mod, par, par.name, type)
{
  nd <- data.frame(nd.mean[,names(nd.mean) != par],
                   par=seq(min(site.stacked[,par]), max(site.stacked[,par]), length.out=1000))
  names(nd) <- gsub(names(nd), pattern="par", replacement=par)
  marginal <- unmarked::predict(mod, type=type, newdata=nd)

  o <- ggplot(marginal, aes(x=nd[,par], y=Predicted)) +
    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.5, fill="grey75") +
    geom_rug(data=site.stacked, aes(x=site.stacked[,par], y=0)) +
    ylim(0,1) +
    xlab(par.name) + ylab(paste("Predicted probability of", ylabel[ylabel$type == type,]$label)) +
    geom_line() + theme_bw(base_size=20) + theme(panel.grid=element_blank())
  print(o)
  ggsave(o, filename=paste("../Figures/occupancy/", sp, "_stacked_occupancy_",type,"_", par,".tiff", sep=""), height=6, width=6, unit="in", dpi=300, compression="lzw")
}

```

## load data
data:
- removed ME 2nd and 3rd surveys (only kept 1st survey to stay consistent with other states)
- removed minutes 7-10 of CT surveys to stay consistent with protocols from other states

```{r}
ewpw.data <- read.csv("../Data_processed/occupancy_detectionhistory_stacked_EWPW.csv")
cwwi.data <- read.csv("../Data_processed/occupancy_detectionhistory_stacked_CWWI.csv")
coni.data <- read.csv("../Data_processed/occupancy_detectionhistory_stacked_CONI.csv")

#select data here
# data <- ewpw.data
data <- cwwi.data
# data <- coni.data

sp <- "CWWI"

# fix moon data
# ME: 1st Sunset (A), 2nd Sunset (B), 1st Moonrise (C), or 2nd Moonrise (D): this is more about start_time
data$moon <- toupper(data$moon)
data$moon <- trimws(data$moon)
data[data$moon %in% c("A","B","C","D"),]$moon <- NA
data[data$moon %in% c("Y"),]$moon <- 1
data[data$moon %in% c("N"),]$moon <- 0
data[data$moon %in% c(""),]$moon <- NA

#remove rows where there is no detection information
data <- data[!is.na(data$MaxDetections),]
data <- data[data$MaxDetections != "-Inf",]

#reorder data by state and year
data <- data[order(data$state, data$year),]

#start with only data with all covariates (remove rows with NA in siteCovs)
summary(data)


#####
#####
#for detection only models and global models
# data <- data[!is.na(data$sky),]
# data <- data[!is.na(data$moon),]
# data <- data[!is.na(data$noise),]
# 
# data <- data[!is.na(data$start_date_j),] #removes most of NH sites since they don't have month/day or start_time info
# data <- data[!data$start_month %in% c(9, 10, 11),] #1 survey in september, 20 surveys in october, and 31 surveys in november

#####
#for occupancy only models and global models
#remove data without BCR information
data <- data[!is.na(data$BCR),] #remove sites without BCR info
data <- data[!is.na(data$year),] #remove sites without year info

data <- data[!is.na(data$Route.Latitude),] #try this to see if latitude is a better predictor
data <- data[!is.na(data$Route.Longitude),]

#remove 2023 data since only from DE
data <- data[data$year != 2023,]

#fix classes for data
data$noise <- as.numeric(data$noise)
data$moon <- factor(data$moon)

data$hour <- factor(data$hour)
data$BCR <- factor(data$BCR)
data$year <- factor(data$year)

######
plot(data$hour, data$TotalDetections)
plot(data$hour, data$MaxDetections)
table(data$state)

nrow(data) #CWWI: 16,561 rows for detection models, 19,737 rows for occupancy models
summary(data)

sum(data$TotalDetections) #this should match number of 1s in the summary of the umf

```

## create umf for stacked single-season occupancy model
```{r}
sp.stacked <- data[,grep(names(data), pattern="min")]
sp.stacked[sp.stacked > 1 & !is.na(sp.stacked)] <- 1 #convert > 1 to 1 (presence/absence instead of counts)

#check detection histories
unique(sp.stacked)

site.stacked <- data[,!names(data) %in% grep(names(data), pattern="min", value=T)] #siteCovs

#add in transformations for variables
site.stacked$start_date_j2 <- scale(site.stacked$start_date_j)*scale(site.stacked$start_date_j)
site.stacked$Route.Latitude2 <- scale(site.stacked$Route.Latitude)^2
site.stacked$Route.Longitude2 <- scale(site.stacked$Route.Longitude)^2

min.df <- matrix(rep(1:6, times=nrow(sp.stacked)), byrow=T, ncol=6) #add obsCovs, dimensions the same as y data

umf <- unmarkedFrameOccu(y=sp.stacked, siteCovs=site.stacked, obsCovs=list(min=min.df))
head(umf)

# tiff(filename="../Figures/occupancy/CWWI_dethist_nsn_stacked.tiff", height=10, width=8, units="in", compression="lzw", res=400)
# plot(umf)
# dev.off()

plot(umf)
summary(umf)

#y=1/(y=0 + y=1)
29891/(51541 + 29891) #only 30.1% of occasions where CWWI were detected (with all NA data removed) for the detection only model
14782/(58275 + 14782) #only 20.2% of occasions where CWWI were detected for the occupancy only model

```

## run detection models for stacked single season occupancy models
include random effect of route and state

13520 sites
Maximum number of observations per site: 6 
Mean number of observations per site: 6 
Sites with at least one detection: 7723 

Tabulation of y observations:
    0     1 
51390 29730
```{r, echo=T}
null <- occu(~1 ~1, data=umf)
null.r <- occu(~(1|st_id) + (1|state) ~1, data=umf)
summary(null.r)


plogis(-5.72) #detection -> 0.3% detection
plogis(1.18)  #occupancy -> 76.5% occupancy

#####
#detection models
#####
p.min <- occu(~scale(min) + (1|st_id) + (1|state) ~1, data=umf)

#
p.moon <- occu(~moon + (1|st_id) + (1|state) ~1, data=umf)
p.sky <- occu(~scale(sky) + (1|st_id) + (1|state) ~1, data=umf)
p.noise <- occu(~scale(noise) + (1|st_id) + (1|state) ~1, data=umf)

p.day <- occu(~scale(start_date_j) + (1|st_id) + (1|state) ~1, data=umf)
p.day2 <- occu(~scale(start_date_j) + I(scale(start_date_j) * scale(start_date_j)) + 
                 (1|st_id) + (1|state) ~1, data=umf)
p.month <- occu(~scale(start_month) + (1|st_id) + (1|state) ~1, data=umf)
p.month2 <- occu(~scale(start_month) + I(scale(start_month)*scale(start_month)) + 
                   (1|st_id) + (1|state) ~1, data=umf)
p.hour <- occu(~hour + (1|st_id) + (1|state) ~1, data=umf)
p.time <- occu(~scale(start_time_s) + (1|st_id) + (1|state) ~1, data=umf)

p.timecat <- occu(~start_time_cat + (1|st_id) + (1|state) ~1, data=umf)

p.global <- occu(~scale(min) + scale(noise) + 
                   scale(start_date_j) + start_date_j2 +
                   start_time_cat + 
                   (1|st_id) + (1|state) ~1, data=umf)

p.global2 <- occu(~scale(min) + scale(noise) + 
                   scale(start_date_j) + I(scale(start_date_j)*scale(start_date_j)) +
                   start_time_cat + 
                   (1|st_id) + (1|state) ~1, data=umf)

p.fl <- fitList(null, null.r, p.min, #per minute within a survey
                p.moon, p.sky, p.noise, #site covariates
                p.day, p.day2, p.month, p.month2, p.time, p.timecat, p.hour) #, #timing of survey
                # p.global) 
modSel(p.fl)
#           nPars      AIC   delta    AICwt cumltvWt
# p.global      8 72694.59    0.00  1.0e+00     1.00
# p.min         3 72821.55  126.96  2.7e-28     1.00
# p.noise       3 72962.21  267.62  7.7e-59     1.00
# p.day2        4 73021.81  327.22  8.8e-72     1.00
# p.month       3 73022.70  328.11  5.6e-72     1.00 #correlated with day2
# p.month2      4 73024.49  329.90  2.3e-72     1.00 #correlated with day2
# p.timecat     5 73028.90  334.31  2.5e-73     1.00

# null.r        2 73030.69  336.10  1.0e-73     1.00
# p.time        3 73032.49  337.90  4.2e-74     1.00
# p.day         3 73051.27  356.68  3.5e-78     1.00
# p.moon        3 73103.00  408.41  2.1e-89     1.00
# p.hour       20 73141.32  446.73  9.8e-98     1.00
# p.sky         3 73151.10  456.51 7.4e-100     1.00
# null          2 79357.23 6662.64  0.0e+00     1.00
```
## coefficients and marginal plots for detection variables
include random effect of route and state
```{r, echo=T}
# mod <- p.global #choose model to extract coefficients from

p.coefs <- vals.coefs(p.global)

c.plot <- ggplot(data=p.coefs[!p.coefs$name %in% "Int",], aes(x=name, y=est, group=p, shape=p)) +
  # c.plot <- ggplot(data=coefs[coefs$name != "INT",], aes(x=name, y=est)) +
  geom_point(size=3, position=position_dodge(width=0.5)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.5)) +
  # geom_errorbar(aes(ymin=est-se, ymax=est+se), width=0, lwd=1, position=position_dodge(width=0.5)) +
  geom_hline(aes(yintercept=0), lty="dotted") +
  xlab("parameter") + ylab("estimate") +
  coord_flip() +
  theme_bw(base_size=16) + theme(panel.grid=element_blank()) + facet_wrap(.~p, scales="free", ncol=1)
c.plot

ggsave(c.plot, filename="../Figures/occupancy/CWWI_stacked_occupancy_det_coefficients.tiff")

###########
#univariate model marginal plots
plotEffects(p.min, "det", covariate="min")
plotEffects(p.noise, "det", covariate="noise")
plotEffects(p.day2, "det", covariate="start_date_j")
plotEffects(p.month2, "det", covariate="start_month")
plotEffects(p.timecat, "det", covariate="start_time_cat")
plotEffects(p.hour, "det", covariate="hour")


###########
# #global model marginal plots for each covariate

############
#plot with base R functions
# #problem seems to be using I(date^2) notation
plotEffects(p.global2, "det", covariate="start_date_j") #switched to start_date_j2
plotEffects(p.global, "det", covariate="min")
plotEffects(p.global, "det", covariate="noise")
# 
# ###########
# #manually plot marginals in ggplot2
nd.mean <- site.stacked %>% mutate_if(is.numeric, mean)
nd.mean <- nd.mean[12,] #only take one row since all rows have the same numeric values (but st_id matters)
nd.mean$min <- 1

ylabel <- data.frame(type=c("det","state"), label=c("detection","occupancy"))
#
#plot these with p.global2, need to pull out scale() from within I() to plot these
plot.marginal(sp="CWWI", p.global, "noise", "noise", type="det")
plot.marginal(sp="CWWI", p.global2, "start_date_j", "Julian day", type="det") #need to plot separately with p.global2 since have scale within I()

#plot separately for categorical and obsCovs
#for categorical
sp <- "CWWI"; par <- "start_time_cat"; par.name <- "start time"; type <- "det"
nd <- data.frame(nd.mean[,names(nd.mean) != par], par=c("day","dawn","night","dusk"))
names(nd) <- gsub(names(nd), pattern="par", replacement=par)
marginal <- predict(p.global, type=type, newdata=nd)
o <- ggplot(marginal, aes(x=nd[,par], y=Predicted)) +
  # geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.5, fill="grey75") + geom_line() + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.5, lwd=1) + geom_point(size=3) +
  ylab(paste("Predicted probability of", ylabel[ylabel$type == type,]$label)) + #ylim(0,1) +
  xlab(par.name) +
  theme_bw(base_size=20) + theme(panel.grid=element_blank())
print(o)
ggsave(o, filename=paste("../Figures/occupancy/", sp, "_stacked_occupancy_",type,"_", par,".tiff", sep=""), height=6, width=6, unit="in", dpi=300, compression="lzw")

#for min
sp <- "CWWI"; par <- "min"; par.name <- "minute"; type <- "det"
nd <- data.frame(nd.mean[,names(nd.mean) != par], par=1:6)
names(nd) <- gsub(names(nd), pattern="par", replacement=par)
marginal <- predict(p.global, type=type, newdata=nd)
o <- ggplot(marginal, aes(x=nd[,par], y=Predicted)) +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.5, fill="grey75") +
  ylab(paste("Predicted probability of", ylabel[ylabel$type == type,]$label)) +ylim(0,1) +
  xlab(par.name) + scale_x_continuous(breaks=1:6) +
  geom_line() + theme_bw(base_size=20) + theme(panel.grid=element_blank())
print(o)
ggsave(o, filename=paste("../Figures/occupancy/", sp, "_stacked_occupancy_",type,"_", par,".tiff", sep=""), height=6, width=6, unit="in", dpi=300, compression="lzw")

```

## run occupancy models for stacked single season occupancy models
include random effect of route on occupancy, included state as a fixed effect
removed sites without BCR information

11567 sites
Maximum number of observations per site: 6 
Mean number of observations per site: 6 
Sites with at least one detection: 3821 

Tabulation of y observations:
    0     1  <NA> 
54609 14782    11

```{r, echo=T}
#####
#occupancy models
#####

null <- occu(~1 ~ (1|st_id), data=umf)

#univariate
o.state <- occu(~1 ~ state + (1|st_id), data=umf)
o.bcr <- occu(~1 ~ BCR + (1|st_id), data=umf) 
o.year <- occu(~1 ~ year + (1|st_id), data=umf)

o.lat <- occu(~1 ~ scale(Route.Latitude) + (1|st_id), data=umf)
o.lat2 <- occu(~1 ~ scale(Route.Latitude) + I(scale(Route.Latitude)* scale(Route.Latitude)) + (1|st_id), data=umf)

o.long <- occu(~1 ~ scale(Route.Longitude) + (1|st_id), data=umf)
o.long2 <- occu(~1 ~ scale(Route.Longitude) + I(scale(Route.Longitude)* scale(Route.Longitude)) + (1|st_id), data=umf)

#univariate model table
o.fl <- fitList(null,
                o.lat, o.lat2, o.long, o.long2,
                o.bcr, o.state, o.year)
modSel(o.fl)
#         nPars      AIC  delta   AICwt cumltvWt
# o.lat2      4 33835.08   0.00 1.0e+00     1.00
# o.state    11 33852.06  16.97 2.1e-04     1.00
# o.lat       3 33889.79  54.71 1.3e-12     1.00
# o.bcr       8 33893.84  58.76 1.7e-13     1.00
# o.long2     4 33907.15  72.07 2.2e-16     1.00
# o.long      3 33915.02  79.94 4.4e-18     1.00
# null        2 34080.56 245.48 5.0e-54     1.00
# o.year     17 34094.78 259.70 4.0e-57     1.00

#these are all the same model, coded differently to be able to plot easily
o.global <- occu(~1 ~ scale(Route.Latitude) + Route.Latitude2 + 
                   BCR + scale(Route.Longitude) + Route.Longitude2 +
                   (1|st_id), data=umf)
o.global0 <- occu(~1 ~ scale(Route.Latitude) + I(scale(Route.Latitude)* scale(Route.Latitude)) + 
                   BCR + scale(Route.Longitude) + I(scale(Route.Longitude) * scale(Route.Longitude)) +
                   (1|st_id), data=umf)
o.global1 <- occu(~1 ~ scale(Route.Latitude) + I(scale(Route.Latitude)* scale(Route.Latitude)) + 
                   BCR + scale(Route.Longitude) + Route.Longitude2 +
                   (1|st_id), data=umf)
o.global2 <- occu(~1 ~ scale(Route.Latitude) + Route.Latitude2 + 
                   BCR + scale(Route.Longitude) + I(scale(Route.Longitude) * scale(Route.Longitude)) +
                   (1|st_id), data=umf)

# o.state.bcr <- occu(~1 ~ state + BCR + (1|st_id), data=umf) #this will not run
o.global.year <- occu(~1 ~ scale(Route.Latitude) + Route.Latitude2 + 
                   BCR + scale(Route.Longitude) + Route.Longitude2 + year + 
                   (1|st_id), data=umf)


```

## plot occupancy only model figures
```{r}
plotEffects(o.lat2, "state", covariate="Route.Latitude")
plotEffects(o.long, "state", covariate="Route.Longitude")
plotEffects(o.long2, "state", covariate="Route.Longitude")
plotEffects(o.state, "state", covariate="state")
plotEffects(o.bcr, "state", covariate="BCR")
# plotEffects(o.year, "state", covariate="year") #this is silly
# 

plotEffects(o.global1, "state", covariate="Route.Latitude")
plotEffects(o.global2, "state", covariate="Route.Longitude")

##########
nd.mean <- site.stacked %>% mutate_if(is.numeric, mean)
nd.mean <- nd.mean[200,] #only take one row since all rows have the same numeric values (but st_id matters)

ylabel <- data.frame(type=c("det","state"), label=c("detection","occupancy"))
#
#plot these with p.global2, need to pull out scale() from within I() to plot these
plot.marginal(sp="CWWI", o.global1, "Route.Latitude", "Latitude", type="state")
plot.marginal(sp="CWWI", o.global2, "Route.Longitude", "Longitude", type="state")


####
#plot occupancy per state per BCR
####
# nd <- data.frame(state=rep(unique(data$state), times=length(unique(data$year))), 
#                  year=rep(unique(data$year), each=length(unique(data$state))),
#                  st_id="GA--270056") #have to provide this for predict to work



#######
#plot values for each BCR
#######
require(sf)
AF_states <- st_read(dsn="../../GIS", layer="AF_states")
BCR.ST <- read_sf(dsn="../../GIS", layer="BCR-state") #area in mi^2
BCR <- BCR.ST %>% group_by(BCR) %>% summarise(AreaSqMi = sum(AreaSqMi, na.rm=T), nstates=n())

mod <- o.global0

#melt state boundaries within BCR
#calculate centroid of each BCR, use to predict
#need to provide st_id for each state
nd.centroid <- st_centroid(BCR) %>% st_coordinates()
st_id <- unique(site.stacked[,c("BCR","st_id")]) %>%
  group_by(BCR) %>%
  arrange(st_id) %>%
  filter(row_number()==1)

nd <- cbind(BCR %>% st_drop_geometry, nd.centroid)
names(nd) <- c("BCR","AreaSqMi","nstate","Route.Longitude","Route.Latitude")

nd <- merge(nd, st_id, by=c("BCR"), all.x=T)
nd$BCR <- factor(nd$BCR)

marginal <- predict(mod, type="state", newdata=nd)
m.data <- cbind(nd, marginal)
# m.data <- m.data[order(m.data$state),]
# m.data$state <- factor(m.data$state, levels=rev(c("FL","GA","SC","NC","VA","WV","MD","DE","NJ","PA","NY","CT","MA","VT","NH","ME")))
m.data$BCR <- factor(m.data$BCR, levels=rev(c(31,27,29,28,30,13,14)))

ggplot(m.data, aes(x=BCR, y=Predicted)) +
  geom_hline(aes(yintercept=0), lty="dashed", col="grey50") + geom_hline(aes(yintercept=1), lty="dashed", col="grey50") +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.25)) +
  geom_point(size=5, col="black", position=position_dodge(width=0.25)) +
  xlab("BCR") + ylab(paste("Predicted probability of occupancy")) + ylim(c(0,1)) +
  theme_bw(base_size=20) + theme(panel.grid=element_blank())
ggsave(filename="../Figures/occupancy/CWWI_stacked_occupancy_psi_BCR.tiff", height=6, width=8, units="in", dpi=400, compression="lzw")


############
#force model to have year so we can look at trend
mod <- o.global.year

u.year <- unique(site.stacked[,c("BCR","year")])
nd <- merge(nd, u.year, by="BCR", all.x=T)
nd$Route.Latitude2 <- scale(nd$Route.Latitude)^2
nd$Route.Longitude2 <- scale(nd$Route.Longitude)^2

marginal <- predict(mod, type="state", newdata=nd)
m.data <- cbind(nd, marginal)
# m.data <- m.data[order(m.data$state),]
# m.data$state <- factor(m.data$state, levels=rev(c("FL","GA","SC","NC","VA","WV","MD","DE","NJ","PA","NY","CT","MA","VT","NH","ME")))
# m.data$BCR <- factor(m.data$BCR, levels=rev(c(31,27,29,28,30,13,14)))

ggplot(m.data, aes(x=year, y=Predicted, col=BCR)) +
  geom_hline(aes(yintercept=0), lty="dashed", col="grey50") + geom_hline(aes(yintercept=1), lty="dashed", col="grey50") +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.75)) +
  geom_point(size=3, position=position_dodge(width=0.75)) +
  scale_color_brewer(palette="Set2") + 
  xlab("Year") + ylab(paste("Predicted probability of occupancy")) + ylim(c(0,1)) +
  theme_bw(base_size=20) + theme(panel.grid=element_blank())
ggsave(filename="../Figures/occupancy/CWWI_stacked_occupancy_psi_BCR-year.tiff", height=6, width=16, units="in", dpi=400, compression="lzw")



#######
#summarize and plot beta coefficients
#this isn't very useful. more useful to use predict and plot those values (as done above)
# psi.glob.info <- vals.coefs(mod = o.state.bcr.year)
# 
# psi.glob.info$level <- gsub(psi.glob.info$name, pattern="[0-9]{2,4}", replacement="")
# psi.glob.info[grep(psi.glob.info$level, pattern="state"),]$level <- "state"
# psi.glob.info$name2 <- gsub(gsub(psi.glob.info$name, pattern="state", replacement=""), pattern="year", replacement="")
# 
# c.data <- psi.glob.info
# c.plot <- ggplot(data=c.data[!c.data$name %in% "Int",], aes(x=name2, y=est, group=p, shape=p)) +
#   # c.plot <- ggplot(data=coefs[coefs$name != "INT",], aes(x=name, y=est)) +
#   geom_point(size=3, position=position_dodge(width=0.5)) + 
#   # geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.5)) +
#   # geom_errorbar(aes(ymin=est-se, ymax=est+se), width=0, lwd=1, position=position_dodge(width=0.5)) +
#   geom_hline(aes(yintercept=0), lty="dotted") +
#   xlab("parameter") + ylab("estimate") +
#   coord_flip() +
#   theme_bw(base_size=16) + theme(panel.grid=element_blank(), legend.position="none") + facet_wrap(.~level, scales="free", ncol=1)

```
## run det + psi models for stacked single season occupancy models
include random effect of route and state on detection
include random effect of route on occupancy

7493 sites
Maximum number of observations per site: 6 
Mean number of observations per site: 6 
Sites with at least one detection: 2359 

Tabulation of y observations:
    0     1 
35833  9125

*this is a much smaller dataset since need covariate info and BCR info*

```{r, echo=T}
#####
#detection + occupancy models
#####
null <- occu(~1 ~1, data=umf)
null.r <- occu(~(1|st_id) + (1|state) ~(1|st_id), data=umf)
  
p.global <- occu(~scale(min) + scale(noise) + 
                   scale(start_date_j) + I(scale(start_date_j)*scale(start_date_j)) +
                   scale(moon) + #scale(sky) + #removed since almost no effect
                   (1|st_id) + (1|state) ~1, data=umf)
o.global <- occu(~1 ~ state + BCR + year + (1|st_id), data=umf)

global <- occu(~scale(min) + scale(noise) + 
                   scale(start_date_j) + I(scale(start_date_j)*scale(start_date_j)) +
                   scale(moon) + #scale(sky) + #remove since almost no effect
                   (1|st_id) + (1|state)
               ~state + BCR + year + (1|st_id), data=umf)

fl <- fitList(null, null.r,
                p.global, o.global, global)
modSel(fl)
```


## plot marginals for stacked single season occupancy models
```{r, echo=T}


# plotEffects(p.month2, "det", covariate="state")
# plotEffects(p.month2, "det", covariate="start_month")


# plotEffects(p.year.min.o.state.year, "state", covariate="state")
# plotEffects(p.year.min.o.state.year, "state", covariate="year")

```