---
title: "6_nightjar surveys_single-season occupancy_stacked years_CONI"
author: "Marie I. Tosa"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(unmarked)
require(plyr)
require(dplyr)
require(tidyr)

require(ggplot2)
require(ggpubr)
```

## Create functions to plot figures
```{r}
vals.coefs <- function(mod)
{
coefs <- coef(mod) %>% tibble::enframe() #extract coefficients for covariates
se <- SE(mod) %>% tibble::enframe() #extract SEs for covariates

ci.psi <- confint(mod, type="state") %>% tibble::enframe()
ci.det <- confint(mod, type="det") %>% tibble::enframe()
ci <- rbind(ci.psi, ci.det) #combine psi and det CIs
ci <- data.frame(name=ci$name, lower=ci$value[,1], upper=ci$value[,2]) #need to split up ci so can read as data frame with lower and upper bounds

coefs <- merge(coefs, se, by="name") #merge coefficient values and ses
coefs <- merge(coefs, ci) #merge coefs with CIs
coefs <- data.frame(coefs)
names(coefs) <- c("par", "est", "se", "lower","upper")

coefs$p <- gsub(substr(coefs$par, 0,3), pattern="p\\([A-z]", replacement="det")
coefs$name <- gsub(coefs$par, pattern="scale\\(", replacement="")
coefs$name <- gsub(coefs$name, pattern="p\\(", replacement="")
coefs$name <- gsub(coefs$name, pattern="psi\\(", replacement="")
coefs$name <- gsub(coefs$name, pattern="\\)\\)", replacement="")
coefs$name <- gsub(coefs$name, pattern="Int\\)", replacement="Int")
coefs$name <- gsub(coefs$name, pattern="\\)", replacement="")
coefs$name <- gsub(coefs$name, pattern="I\\(", replacement="")
return(coefs)
}

plot.marginal <- function(sp, mod, par, par.name, type)
{
  nd <- data.frame(nd.mean[,names(nd.mean) != par],
                   par=seq(min(site.stacked[,par]), max(site.stacked[,par]), length.out=1000))
  names(nd) <- gsub(names(nd), pattern="par", replacement=par)
  marginal <- unmarked::predict(mod, type=type, newdata=nd)

  o <- ggplot(marginal, aes(x=nd[,par], y=Predicted)) +
    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.5, fill="grey75") +
    geom_rug(data=site.stacked, aes(x=site.stacked[,par], y=0)) +
    ylim(0,1) +
    xlab(par.name) + ylab(paste("Predicted probability of", ylabel[ylabel$type == type,]$label)) +
    geom_line() + theme_bw(base_size=20) + theme(panel.grid=element_blank())
  print(o)
  ggsave(o, filename=paste("../Figures/occupancy/", sp, "_stacked_occupancy_",type,"_", par,".tiff", sep=""), height=6, width=6, unit="in", dpi=300, compression="lzw")
}

```

## load data
data:
- removed ME 2nd and 3rd surveys (only kept 1st survey to stay consistent with other states)
- removed minutes 7-10 of CT surveys to stay consistent with protocols from other states

```{r}
ewpw.data <- read.csv("../Data_processed/occupancy_detectionhistory_stacked_EWPW.csv")
cwwi.data <- read.csv("../Data_processed/occupancy_detectionhistory_stacked_CWWI.csv")
coni.data <- read.csv("../Data_processed/occupancy_detectionhistory_stacked_CONI.csv")

#select data here
# data <- ewpw.data
# data <- cwwi.data
data <- coni.data
sp <- "CONI"

# fix moon data
# ME: 1st Sunset (A), 2nd Sunset (B), 1st Moonrise (C), or 2nd Moonrise (D): this is more about start_time
data$moon <- toupper(data$moon)
data$moon <- trimws(data$moon)
data[data$moon %in% c("A","B","C","D"),]$moon <- NA
data[data$moon %in% c("Y"),]$moon <- 1
data[data$moon %in% c("N"),]$moon <- 0
data[data$moon %in% c(""),]$moon <- NA

#remove rows where there is no detection information
data <- data[!is.na(data$MaxDetections),]
data <- data[data$MaxDetections != "-Inf",]

#reorder data by state and year
data <- data[order(data$state, data$year),]

#start with only data with all covariates (remove rows with NA in siteCovs)
summary(data)


#####
#for detection only models and global models
# data <- data[!is.na(data$sky),]
# data <- data[!is.na(data$moon),]
# data <- data[!is.na(data$noise),]
# 
# data <- data[!is.na(data$start_date_j),] #removes most of NH sites since they don't have month/day or start_time info
# # data <- data[!data$start_month %in% c(9, 10, 11),]

#####
#for occupancy only models and global models
#remove data without BCR information
data <- data[!is.na(data$BCR),] #remove sites without BCR info
data <- data[!is.na(data$year),] #remove sites without year info

data <- data[!is.na(data$Route.Latitude),] #try this to see if latitude is a better predictor
data <- data[!is.na(data$Route.Longitude),]

#remove 2023 data since only from DE
data <- data[data$year != 2023,]

#fix classes for data
data$noise <- as.numeric(data$noise)
# data$moon <- factor(data$moon)
data$moon <- as.numeric(data$moon)

data$hour <- factor(data$hour)
data$start_month <- factor(data$start_month)
data$BCR <- factor(data$BCR)
data$year <- factor(data$year)

######
plot(data$hour, data$TotalDetections)
plot(data$hour, data$MaxDetections)
table(data$state)

nrow(data) #CWWI: 16,561 rows for detection models, 19,737 rows for occupancy models
summary(data)

sum(data$TotalDetections) #this should match number of 1s in the summary of the umf

```

## create umf for stacked single-season occupancy model
```{r}
sp.stacked <- data[,grep(names(data), pattern="min")]
sp.stacked[sp.stacked > 1 & !is.na(sp.stacked)] <- 1 #convert > 1 to 1 (presence/absence instead of counts)

#check detection histories
unique(sp.stacked)

site.stacked <- data[,!names(data) %in% grep(names(data), pattern="min", value=T)] #siteCovs

#add in transformations for variables
site.stacked$start_date_j2 <- scale(site.stacked$start_date_j)*scale(site.stacked$start_date_j)
site.stacked$start_time2 <- scale(site.stacked$start_time_s)*scale(site.stacked$start_time_s)

site.stacked$Route.Latitude2 <- scale(site.stacked$Route.Latitude)^2
site.stacked$Route.Longitude2 <- scale(site.stacked$Route.Longitude)^2

min.df <- matrix(rep(1:6, times=nrow(sp.stacked)), byrow=T, ncol=6) #add obsCovs, dimensions the same as y data

umf <- unmarkedFrameOccu(y=sp.stacked, siteCovs=site.stacked, obsCovs=list(min=min.df))
head(umf)

tiff(filename="../Figures/occupancy/CONI_dethist_nsn_stacked.tiff", height=10, width=8, units="in", compression="lzw", res=400)
plot(umf)
dev.off()

plot(umf)
summary(umf)

#y=1/(y=0 + y=1)
1822 /(45434 + 1822) #only 3.9% of occasions where CONI were detected (with all NA data removed) for the detection only model
 #only 3.5% of occasions where CONI were detected for the occupancy only model

```

## run detection models for stacked single season occupancy models
include random effect of route and state

7876 sites
Maximum number of observations per site: 6 
Mean number of observations per site: 6 
Sites with at least one detection: 567 

Tabulation of y observations:
    0     1 
45434  1822

```{r, echo=T}
null <- occu(~1 ~1, data=umf)
null.r <- occu(~(1|st_id) + (1|state) ~1, data=umf)
summary(null.r)

plogis(-5.86) #detection -> 0.2% detection
plogis(-1.26)  #occupancy -> 22.1% occupancy

#####
#detection models
#####
p.min <- occu(~scale(min) + (1|st_id) + (1|state) ~1, data=umf)

#
# p.moon <- occu(~moon + (1|st_id) + (1|state) ~1, data=umf) #NaNs produced
p.sky <- occu(~scale(sky) + (1|st_id) + (1|state) ~1, data=umf)
p.noise <- occu(~scale(noise) + (1|st_id) + (1|state) ~1, data=umf)

p.day <- occu(~scale(start_date_j) + (1|st_id) + (1|state) ~1, data=umf)
p.day2 <- occu(~scale(start_date_j) + I(scale(start_date_j) * scale(start_date_j)) + 
                 (1|st_id) + (1|state) ~1, data=umf)
# p.month <- occu(~start_month + (1|st_id) + (1|state) ~1, data=umf) #wild SEs
# p.month2 <- occu(~scale(start_month) + I(scale(start_month)*scale(start_month)) + 
                   # (1|st_id) + (1|state) ~1, data=umf)
# p.hour <- occu(~hour + (1|st_id) + (1|state) ~1, data=umf) #wild SEs
p.time <- occu(~scale(start_time_s) + (1|st_id) + (1|state) ~1, data=umf)
p.time2 <- occu(~scale(start_time_s) + I(scale(start_time_s)*scale(start_time_s)) + (1|st_id) + (1|state) ~1, data=umf)

p.timecat <- occu(~start_time_cat + (1|st_id) + (1|state) ~1, data=umf)

####
p.fl <- fitList(null, null.r, p.min, #per minute within a survey
                # p.moon, 
                p.sky, p.noise, #site covariates
                p.day, p.day2, p.time, p.time2, p.timecat) #p.hour, p.month #, #timing of survey p.month2, 
                # p.global) 
modSel(p.fl)
#            nPars     AIC   delta    AICwt cumltvWt
# p.time2       4 7763.62    0.00  8.1e-01     0.81
# p.noise       3 7766.88    3.26  1.6e-01     0.96
# p.timecat     5 7771.71    8.09  1.4e-02     0.98 #
# p.day         3 7772.27    8.65  1.1e-02     0.99
# p.day2        4 7773.47    9.85  5.9e-03     0.99
# p.min         3 7775.69   12.07  1.9e-03     1.00
# p.sky         3 7776.19   12.56  1.5e-03     1.00

# null.r        2 7776.43   12.80  1.3e-03     1.00
# p.time        3 7777.75   14.13  6.9e-04     1.00
# null          2 8767.04 1003.42 1.0e-218     1.00

p.global <- occu(~scale(start_time_s) + start_time2 + scale(noise) + scale(sky) + scale(start_date_j) + scale(min) + moon +
                   (1|st_id) + (1|state) ~1, data=umf)
#same as above but with scale(original variable)
p.global2 <- occu(~scale(start_time_s) + I(scale(start_time_s)^2) + scale(noise) + scale(sky) + scale(start_date_j) + scale(min) + moon +
                   (1|st_id) + (1|state) ~1, data=umf)
```
## coefficients and marginal plots for detection variables
include random effect of route and state
```{r, echo=T}
# mod <- p.global #choose model to extract coefficients from

p.coefs <- vals.coefs(p.global)

c.plot <- ggplot(data=p.coefs[!p.coefs$name %in% "Int",], aes(x=name, y=est, group=p, shape=p)) +
  # c.plot <- ggplot(data=coefs[coefs$name != "INT",], aes(x=name, y=est)) +
  geom_point(size=3, position=position_dodge(width=0.5)) + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.5)) +
  # geom_errorbar(aes(ymin=est-se, ymax=est+se), width=0, lwd=1, position=position_dodge(width=0.5)) +
  geom_hline(aes(yintercept=0), lty="dotted") +
  xlab("parameter") + ylab("estimate") +
  coord_flip() +
  theme_bw(base_size=16) + theme(panel.grid=element_blank(), legend.position="none") + facet_wrap(.~p, scales="free", ncol=1)
c.plot

ggsave(c.plot, filename="../Figures/occupancy/CONI_stacked_occupancy_det_coefficients.tiff", height=6, width=4, units="in", dpi=400, compression="lzw")

###########
#univariate model marginal plots
plotEffects(p.min, "det", covariate="min")
plotEffects(p.noise, "det", covariate="noise")
plotEffects(p.day2, "det", covariate="start_date_j")
# plotEffects(p.month2, "det", covariate="start_month")
plotEffects(p.timecat, "det", covariate="start_time_cat")
plotEffects(p.hour, "det", covariate="hour")


###########
# #global model marginal plots for each covariate

############
#plot with base R functions
# #problem seems to be using I(date^2) notation
plotEffects(p.global, "det", covariate="start_date_j") #switched to start_date_j2
plotEffects(p.global, "det", covariate="min")
plotEffects(p.global, "det", covariate="noise")
# 
# ###########
# #manually plot marginals in ggplot2
nd.mean <- site.stacked %>% mutate_if(is.numeric, mean)
nd.mean <- nd.mean[12,] #only take one row since all rows have the same numeric values (but st_id matters)
nd.mean$min <- 1

ylabel <- data.frame(type=c("det","state"), label=c("detection","occupancy"))
#
#plot these with p.global2, need to pull out scale() from within I() to plot these
plot.marginal(sp="CONI", p.global, "noise", "noise", type="det")
plot.marginal(sp="CONI", p.global, "start_date_j", "Julian day", type="det")
plot.marginal(sp="CONI", p.global2, "start_time_s", "start time", type="det")
plot.marginal(sp="CONI", p.global, "sky", "sky", type="det")

#plot separately for quadratic, categorical, and obsCovs
#moon
sp <- "CONI"; par <- "moon"; par.name <- "moon visible"; type <- "det"
nd <- data.frame(nd.mean[,names(nd.mean) != par], par=c("0","1"))
names(nd) <- gsub(names(nd), pattern="par", replacement=par)
marginal <- predict(p.global, type=type, newdata=nd)
o <- ggplot(marginal, aes(x=nd[,par], y=Predicted)) +
  # geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.5, fill="grey75") + geom_line() + 
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.5, lwd=1) + geom_point(size=3) +
  ylab(paste("Predicted probability of", ylabel[ylabel$type == type,]$label)) + #ylim(0,1) +
  xlab(par.name) +
  theme_bw(base_size=20) + theme(panel.grid=element_blank())
print(o)
ggsave(o, filename=paste("../Figures/occupancy/", sp, "_stacked_occupancy_",type,"_", par,".tiff", sep=""), height=6, width=6, unit="in", dpi=300, compression="lzw")

#min
sp <- "CONI"; par <- "min"; par.name <- "minute"; type <- "det"
nd <- data.frame(nd.mean[,names(nd.mean) != par], par=1:6)
names(nd) <- gsub(names(nd), pattern="par", replacement=par)
marginal <- predict(p.global, type=type, newdata=nd)
o <- ggplot(marginal, aes(x=nd[,par], y=Predicted)) +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.5, fill="grey75") + geom_line() +
  # geom_errorbar(aes(ymin=lower, ymax=upper), width=0.5, lwd=1) + geom_point(size=3) +
  ylab(paste("Predicted probability of", ylabel[ylabel$type == type,]$label)) + ylim(0,1) +
  xlab(par.name) + scale_x_continuous(breaks=1:6) +
  theme_bw(base_size=20) + theme(panel.grid=element_blank())
print(o)
ggsave(o, filename=paste("../Figures/occupancy/", sp, "_stacked_occupancy_",type,"_", par,".tiff", sep=""), height=6, width=6, unit="in", dpi=300, compression="lzw")


```

## run occupancy models for stacked single season occupancy models
include random effect of route on occupancy, included state as a fixed effect
removed sites without BCR information

11745 sites
Maximum number of observations per site: 6 
Mean number of observations per site: 6 
Sites with at least one detection: 872 

Tabulation of y observations:
    0     1  <NA> 
67803  2656    11 

  FL   GA   MD   ME   NC   NH   NY   PA   SC   VA   WV 
2145  544  116 3061  809 1844 1514  356  531  813   12 

MaxDetections
state    0    1    2    3    4    5    7    9   24
   FL 1708  296  108   27    5    1    0    0    0
   GA  535    9    0    0    0    0    0    0    0
   MD   92   23    1    0    0    0    0    0    0
   ME 2778  211   49   15    5    0    1    1    1
   NC  793   12    4    0    0    0    0    0    0
   NH 1844    0    0    0    0    0    0    0    0 ** is this real?
   NY 1493   21    0    0    0    0    0    0    0
   PA  351    5    0    0    0    0    0    0    0
   SC  468   37   14    8    3    1    0    0    0
   VA  800   11    1    0    1    0    0    0    0
   WV   11    1    0    0    0    0    0    0    0
   
```{r, echo=T}
#####
#occupancy models
#####

####### did NH actually survey for CONI?? 0 detections in NH

null <- occu(~1 ~ (1|st_id), data=umf)

#univariate
o.state <- occu(~1 ~ state + (1|st_id), data=umf)
o.bcr <- occu(~1 ~ BCR + (1|st_id), data=umf) 
o.year <- occu(~1 ~ year + (1|st_id), data=umf)

o.lat <- occu(~1 ~ scale(Route.Latitude) + (1|st_id), data=umf)
o.lat2 <- occu(~1 ~ scale(Route.Latitude) + I(scale(Route.Latitude)* scale(Route.Latitude)) + (1|st_id), data=umf)

o.long <- occu(~1 ~ scale(Route.Longitude) + (1|st_id), data=umf)
o.long2 <- occu(~1 ~ scale(Route.Longitude) + I(scale(Route.Longitude)* scale(Route.Longitude)) + (1|st_id), data=umf)

#univariate model table
o.fl <- fitList(o.lat, o.lat2, o.long, o.long2)
modSel(o.fl)
#         nPars      AIC delta   AICwt cumltvWt
# o.lat2      4 11899.57  0.00 1.0e+00     1.00
# o.long2     4 11944.12 44.55 2.1e-10     1.00
# o.lat       3 11951.92 52.36 4.3e-12     1.00
# o.long      3 11965.19 65.62 5.6e-15     1.00

o.fl <- fitList(null, o.bcr, o.state, o.year,
                o.lat, o.lat2, o.long, o.long2)
modSel(o.fl)
#         nPars      AIC  delta   AICwt cumltvWt
# o.state    12 11854.33   0.00 1.0e+00     1.00
# o.lat2      4 11899.57  45.24 1.5e-10     1.00
# o.bcr       8 11911.47  57.14 3.9e-13     1.00
# o.long2     4 11944.12  89.79 3.2e-20     1.00
# o.lat       3 11951.92  97.60 6.4e-22     1.00

# null        2 11963.34 109.02 2.1e-24     1.00
# o.long      3 11965.19 110.86 8.5e-25     1.00
# o.year     17 11966.51 112.19 4.4e-25     1.00

o.state.year <- occu(~1 ~ state + year + (1|st_id), data=umf)
o.bcr.year <- occu(~1 ~ BCR + year + (1|st_id), data=umf)

o.state.bcr.year <- occu(~1 ~ state + BCR + year + (1|st_id), data=umf) #NH SE values are crazy because no detections in NH

o.global <- occu(~1 ~ scale(Route.Latitude) + Route.Latitude2 + 
                   scale(Route.Longitude) + Route.Longitude2 +
                   BCR + year + (1|st_id), data=umf)
o.global1 <- occu(~1 ~ scale(Route.Latitude) + I(scale(Route.Latitude)^2) + 
                   scale(Route.Longitude) + Route.Longitude2 +
                   BCR + year + (1|st_id), data=umf)
o.global2 <- occu(~1 ~ scale(Route.Latitude) + Route.Latitude2 + 
                   scale(Route.Longitude) + I(scale(Route.Longitude)^2) +
                   BCR + year + (1|st_id), data=umf)

```

## plot occupancy only model figures
```{r}
plotEffects(o.lat2, "state", covariate="Route.Latitude")
plotEffects(o.long2, "state", covariate="Route.Longitude")

plotEffects(o.state, "state", covariate="state")
plotEffects(o.bcr, "state", covariate="BCR")
# plotEffects(o.year, "state", covariate="year") #this is silly

#####
nd.mean <- site.stacked %>% mutate_if(is.numeric, mean)
nd.mean <- nd.mean[12,] #only take one row (plots depend on which site chosen)
nd.mean$min <- 1

ylabel <- data.frame(type=c("det","state"), label=c("detection","occupancy"))

plot.marginal(sp=sp, mod=o.global1, par="Route.Latitude", par.name="Latitude", type="state")
plot.marginal(sp=sp, mod=o.global2, par="Route.Longitude", par.name="Longitude", type="state")
#

require(sf)
AF_states <- st_read(dsn="../../GIS", layer="AF_states")
BCR.ST <- read_sf(dsn="../../GIS", layer="BCR-state") #area in mi^2
BCR <- BCR.ST %>% group_by(BCR) %>% dplyr::summarise(AreaSqMi = sum(AreaSqMi, na.rm=T), nstates=n())

#melt state boundaries within BCR
#calculate centroid of each BCR, use to predict
#need to provide st_id for each state
nd.centroid <- st_centroid(BCR) %>% st_coordinates()
st_id <- unique(site.stacked[,c("BCR","st_id","year")]) %>%
  group_by(BCR, year) %>%
  arrange(st_id) %>%
  filter(row_number()==1)

nd <- cbind(BCR %>% st_drop_geometry, nd.centroid)
names(nd) <- c("BCR","AreaSqMi","nstate","Route.Longitude","Route.Latitude")

nd <- merge(nd, st_id, by=c("BCR"), all.x=T)
nd$BCR <- factor(nd$BCR)

nd$Route.Latitude2 <- scale(nd$Route.Latitude)^2
nd$Route.Longitude2 <- scale(nd$Route.Longitude)^2
# nd$year <- factor(2018)

mod <- o.global
marginal <- predict(mod, type="state", newdata=nd)
m.data <- cbind(nd, marginal)
# m.data <- m.data[order(m.data$state),]
# m.data$state <- factor(m.data$state, levels=rev(c("FL","GA","SC","NC","VA","WV","MD","DE","NJ","PA","NY","CT","MA","VT","NH","ME")))
# m.data$BCR <- factor(m.data$BCR, levels=rev(c(31,27,29,28,30,13,14)))
m.data$BCR <- factor(m.data$BCR) #make sure colors match the main map
m.data$year <- factor(m.data$year)

ggplot(m.data, aes(x=year, y=Predicted, col=BCR)) +
  scale_color_brewer(palette="Set2") +
  geom_hline(aes(yintercept=0), lty="dashed", col="grey50") + geom_hline(aes(yintercept=1), lty="dashed", col="grey50") +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.25)) +
  geom_point(size=5, position=position_dodge(width=0.25)) +
  xlab("BCR") + ylab(paste("Predicted probability of occupancy")) + ylim(c(0,1)) +
  theme_bw(base_size=20) + theme(panel.grid=element_blank())
ggsave(filename="../Figures/occupancy/CONI_stacked_occupancy_psi_year-BCR.tiff", height=6, width=16, units="in", dpi=400, compression="lzw")

#######
#plot values for each BCR
#######
require(sf)
AF_states <- st_read(dsn="../../GIS", layer="AF_states")
BCR.ST <- read_sf(dsn="../../GIS", layer="BCR-state") #area in mi^2
BCR <- BCR.ST %>% group_by(BCR) %>% dplyr::summarise(AreaSqMi = sum(AreaSqMi, na.rm=T), nstates=n())

mod <- o.global

#melt state boundaries within BCR
#calculate centroid of each BCR, use to predict
#need to provide st_id for each state
nd.centroid <- st_centroid(BCR) %>% st_coordinates()
st_id <- unique(site.stacked[,c("BCR","st_id")]) %>%
  group_by(BCR) %>%
  arrange(st_id) %>%
  filter(row_number()==1)

nd <- cbind(BCR %>% st_drop_geometry, nd.centroid)
names(nd) <- c("BCR","AreaSqMi","nstate","Route.Longitude","Route.Latitude")
nd$Route.Latitude2 <- scale(nd$Route.Latitude)^2
nd$Route.Longitude2 <- scale(nd$Route.Longitude)^2
nd$year <- factor(2018)

nd <- merge(nd, st_id, by=c("BCR"), all.x=T)
nd$BCR <- factor(nd$BCR)

marginal <- predict(mod, type="state", newdata=nd)
m.data <- cbind(nd, marginal)
# m.data <- m.data[order(m.data$state),]
# m.data$state <- factor(m.data$state, levels=rev(c("FL","GA","SC","NC","VA","WV","MD","DE","NJ","PA","NY","CT","MA","VT","NH","ME")))
m.data$BCR <- factor(m.data$BCR, levels=c(31,27,29,28,30,13,14))

ggplot(m.data, aes(x=BCR, y=Predicted)) +
  geom_hline(aes(yintercept=0), lty="dashed", col="grey50") + geom_hline(aes(yintercept=1), lty="dashed", col="grey50") +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.25)) +
  geom_point(size=5, col="black", position=position_dodge(width=0.25)) +
  xlab("BCR") + ylab(paste("Predicted probability of occupancy")) + ylim(c(0,1)) +
  theme_bw(base_size=20) + theme(panel.grid=element_blank())
ggsave(filename="../Figures/occupancy/CONI_stacked_occupancy_psi_BCR.tiff", height=6, width=8, units="in", dpi=400, compression="lzw")






#####
#plot occupancy per state per BCR per year
#####
# nd <- data.frame(state=rep(unique(data$state), times=length(unique(data$year))), 
#                  year=rep(unique(data$year), each=length(unique(data$state))),
#                  st_id="GA--270056") #have to provide this for predict to work

mod <- o.state.bcr.year

#this isn't quite right since could be using different routes each year
# nd <- unique(site.stacked[,c("state","BCR","year","st_id")])
# nd <- nd %>% #take first row of each group
#   group_by(state, BCR, year) %>%
#   arrange(st_id) %>%
#   filter(row_number()==1)

nd <- unique(site.stacked[,c("state","BCR","st_id")])
nd <- nd %>% #take first row of each group
  group_by(state, BCR) %>%
  arrange(st_id) %>%
  filter(row_number()==1)

u.year <- unique(site.stacked[,c("state","BCR","year")])
nd <- merge(nd, u.year, by=c("state","BCR"), all=T)

marginal <- predict(mod, type="state", newdata=nd)

m.data <- cbind(nd, marginal)
m.data <- m.data[order(m.data$year, m.data$state),]
m.data$state <- factor(m.data$state, levels=rev(c("FL","GA","SC","NC","VA","WV","MD","DE","NJ","PA","NY","CT","MA","VT","NH","ME")))

p <- ggplot(m.data, aes(x=year, y=Predicted, col=BCR, fill=BCR)) +
  geom_hline(aes(yintercept=0), lty="dashed", col="grey50") + geom_hline(aes(yintercept=1), lty="dashed", col="grey50") +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.25)) +
  geom_point(size=5, pch=21, col="black", position=position_dodge(width=0.25)) +
  # geom_smooth(method="lm", col="grey50", lty="dashed") +
  scale_fill_brewer(palette="Set2") + scale_color_brewer(palette="Set2") +
  xlab("state") + ylab(paste("Predicted probability of occupancy")) + ylim(c(0,1)) +
  theme_bw(base_size=20) + theme(panel.grid=element_blank())
p + facet_wrap(~state, ncol=1)
ggsave(filename = "../Figures/occupancy/CONI_stacked_occupancy_psi_state-bcr-year.tiff", height=30, width=16, units="in", dpi=400, compression="lzw")

p + facet_wrap(~state, ncol=2, dir="v")
ggsave(filename = "../Figures/occupancy/CONI_stacked_occupancy_psi_state-bcr-year2.tiff", height=20, width=20, units="in", dpi=400, compression="lzw")

######
#plot values for each state
######
mod <- o.state
nd <- unique(site.stacked[,c("state","BCR","year","st_id")])
nd <- nd %>% #take first row of each group
  group_by(state) %>%
  arrange(st_id) %>%
  filter(row_number()==1)
marginal <- predict(mod, type="state", newdata=nd)
m.data <- cbind(nd, marginal)
m.data <- m.data[order(m.data$year, m.data$state),]
m.data$state <- factor(m.data$state, levels=rev(c("FL","GA","SC","NC","VA","WV","MD","DE","NJ","PA","NY","CT","MA","VT","NH","ME")))

ggplot(m.data, aes(x=state, y=Predicted)) +
  geom_hline(aes(yintercept=0), lty="dashed", col="grey50") + geom_hline(aes(yintercept=1), lty="dashed", col="grey50") +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.25)) +
  geom_point(size=5, col="black", position=position_dodge(width=0.25)) +
  xlab("state") + ylab(paste("Predicted probability of occupancy")) + ylim(c(0,1)) +
  theme_bw(base_size=20) + theme(panel.grid=element_blank())
ggsave(filename="../Figures/occupancy/CONI_stacked_occ_psi_state.tiff", height=6, width=8, units="in", dpi=400, compression="lzw")

#######
#plot values for each BCR
#######
mod <- o.bcr
nd <- unique(site.stacked[,c("state","BCR","year","st_id")])
nd <- nd %>% #take first row of each group
  group_by(BCR) %>%
  arrange(st_id) %>%
  filter(row_number()==1)
marginal <- predict(mod, type="state", newdata=nd)
m.data <- cbind(nd, marginal)
m.data <- m.data[order(m.data$year, m.data$state),]
m.data$state <- factor(m.data$state, levels=rev(c("FL","GA","SC","NC","VA","WV","MD","DE","NJ","PA","NY","CT","MA","VT","NH","ME")))
m.data$BCR <- factor(m.data$BCR, levels=rev(c(31,27,29,28,30,13,14)))

ggplot(m.data, aes(x=BCR, y=Predicted)) +
  geom_hline(aes(yintercept=0), lty="dashed", col="grey50") + geom_hline(aes(yintercept=1), lty="dashed", col="grey50") +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.25)) +
  geom_point(size=5, col="black", position=position_dodge(width=0.25)) +
  xlab("BCR") + ylab(paste("Predicted probability of occupancy")) + ylim(c(0,1)) +
  theme_bw(base_size=20) + theme(panel.grid=element_blank())
ggsave(filename="../Figures/occupancy/CONI_stacked_occ_psi_BCR.tiff", height=6, width=8, units="in", dpi=400, compression="lzw")

#can't really plot values for each year since each state was only samples sporadically

#######
#summarize and plot beta coefficients
#this isn't very useful. more useful to use predict and plot those values (as done above)
# psi.glob.info <- vals.coefs(mod = o.state.bcr.year)
# 
# psi.glob.info$level <- gsub(psi.glob.info$name, pattern="[0-9]{2,4}", replacement="")
# psi.glob.info[grep(psi.glob.info$level, pattern="state"),]$level <- "state"
# psi.glob.info$name2 <- gsub(gsub(psi.glob.info$name, pattern="state", replacement=""), pattern="year", replacement="")
# 
# c.data <- psi.glob.info
# c.plot <- ggplot(data=c.data[!c.data$name %in% "Int",], aes(x=name2, y=est, group=p, shape=p)) +
#   # c.plot <- ggplot(data=coefs[coefs$name != "INT",], aes(x=name, y=est)) +
#   geom_point(size=3, position=position_dodge(width=0.5)) + 
#   # geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, lwd=1, position=position_dodge(width=0.5)) +
#   # geom_errorbar(aes(ymin=est-se, ymax=est+se), width=0, lwd=1, position=position_dodge(width=0.5)) +
#   geom_hline(aes(yintercept=0), lty="dotted") +
#   xlab("parameter") + ylab("estimate") +
#   coord_flip() +
#   theme_bw(base_size=16) + theme(panel.grid=element_blank(), legend.position="none") + facet_wrap(.~level, scales="free", ncol=1)

```
## run det + psi models for stacked single season occupancy models
include random effect of route and state on detection
include random effect of route on occupancy



*this is a much smaller dataset since need covariate info and BCR info*

```{r, echo=T}
#####
#detection + occupancy models
#####
null <- occu(~1 ~1, data=umf)
null.r <- occu(~(1|st_id) + (1|state) ~(1|st_id), data=umf)
  
p.global <- occu(~scale(min) + scale(noise) + 
                   scale(start_date_j) + I(scale(start_date_j)*scale(start_date_j)) +
                   scale(moon) + #scale(sky) + #removed since almost no effect
                   (1|st_id) + (1|state) ~1, data=umf)
o.global <- occu(~1 ~ state + BCR + year + (1|st_id), data=umf)

global <- occu(~scale(min) + scale(noise) + 
                   scale(start_date_j) + I(scale(start_date_j)*scale(start_date_j)) +
                   scale(moon) + #scale(sky) + #remove since almost no effect
                   (1|st_id) + (1|state)
               ~state + BCR + year + (1|st_id), data=umf)

fl <- fitList(null, null.r,
                p.global, o.global, global)
modSel(fl)
```


## plot marginals for stacked single season occupancy models
```{r, echo=T}


# plotEffects(p.month2, "det", covariate="state")
# plotEffects(p.month2, "det", covariate="start_month")


# plotEffects(p.year.min.o.state.year, "state", covariate="state")
# plotEffects(p.year.min.o.state.year, "state", covariate="year")

```