---
title: "Nightjar exploratory analysis"
author: "Marie I. Tosa"
date: "`r Sys.Date()`"
output: pdf_document
---

## document setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages

#for playing with data
require(tidyr) #for separate
require(readxl)
require(dplyr) #for bind_rows

#for GIS
require(sf)
require(raster)
require(units)

#for plotting
require(ggplot2)


```

## GIS layers

Look at states within the Atlantic Flyway and the Bird Conservation Regions (BCR)

```{r}

AF_states <- st_read(dsn="../../GIS", layer="AF_states") 
# BCR <- read_sf(dsn="../../GIS", layer="BCR_AF_clip") #don't use this one. not up to date. This one was from Tony Roberts
#these are in NAD 83, longlat

BCR.all <- read_sf(dsn="../../GIS/BCR_Terrestrial", layer="BCR_Terrestrial_master") #https://nabci-us.org/resources/bird-conservation-regions-map/
#http://www.birdscanada.org/research/gislab/index.jsp?targetpg=bcr

AF_states.df <- AF_states
st_geometry(AF_states.df) <- NULL
AF_states.df$STATE_NAME <- toupper(AF_states.df$STATE_NAME)

BCR.AF <- BCR.all[BCR.all$PROVINCE_S %in% toupper(AF_states$STATE_NAME),]
BCR.AF <- merge(BCR.AF, AF_states.df[,c("STATE_NAME","ST")], 
                by.x="PROVINCE_S", by.y="STATE_NAME", all.x=T)
BCR.AF <- st_collection_extract(BCR.AF, "POLYGON")
BCR.AF$AreaSqMi <- set_units(st_area(BCR.AF), "mi^2")

# st_write(BCR.AF, "../../GIS/BCR-state.shp", append=F) #to save

BCR <- read_sf(dsn="../../GIS", layer="BCR-state") #Area in m^2

plot(AF_states)
plot(BCR)

unique(BCR$BCR)
#7 BCRs
#14 13 30 28 29 27 31

AF_states$ST
#18 states
# "CT" "DE" "DC" "FL" "GA" "ME" "MD" "MA" "NH" "NJ" "NY" "NC" "PA" "RI"
# "SC" "VT" "VA" "WV"

AF_states$AreaSqMi <- set_units(st_area(AF_states), "mi^2")

#don't need to do this with the new BCR shapefile from 
#create polygon shapefile with each state cut by BCR
# BCR.ST <- st_intersection(AF_states, BCR)
# BCR.ST$BCR.ST <- paste(BCR.ST$ST, BCR.ST$BCRNumber, sep=".")
# BCR.ST <- st_collection_extract(BCR.ST, "POLYGON") #otherwise has polygons, multipolygons, geometrycollection, etc.
# 
# sf_use_s2(FALSE)
# BCR.ST$AreaBS <- set_units(st_area(BCR.ST), "mi^2")

# st_write(BCR.ST, "../../GIS/BCR-state.shp", append=F) #to save
```

## Raw data from Nightjar Survey Network
format data so easier to work with later

```{r, echo=T}
#raw files
nsn <- read.csv(file="../Data_raw/Copy of NSN Data/NSN Data (March 28 2023) - Routes.csv")
#duplicate routeid when there are stops 1-11
  nsn.all <- nsn
  nsn <- nsn[nsn$Stop == 1 | nsn$Routeid == 981234,] #subset for first stop in each route, no Stop 1 for 981234

rt.id <- read.csv("../Data_raw/Copy of NSN Data/NSN Data (March 28 2023) - Route ID.csv")

nsn.sp <- read.csv("../Data_raw/NSN_sp_key.csv")

nsn.time <- read.csv("../Data_raw/Copy of NSN Data/NSN Data (March 28 2023) - Time and Miles.csv")
  nsn.time <- separate(nsn.time, col="start_time", remove=F,
                       into=c("start_date","start_hour"), sep=" ")
  nsn.time <- separate(nsn.time, col="end_time", remove=F,
                       into=c("end_date","end_hour"), sep=" ")
  nsn.time <- separate(nsn.time, col="start_date",
                       into=c("start_year","start_month","start_day"),
                       sep="-", remove=F)
  
  unique(nsn.time[nsn.time$route_id < 10000,]$route_id)
  # unique(nsn.time[nsn.time$route_id < 100000,]$route_id)
  
  #correct these route_ids
  nsn.time[nsn.time$route_id == 6101,]$route_id <- 60101
  nsn.time[nsn.time$route_id == 6051,]$route_id <- 60051

nsn.obs <- read.csv("../Data_raw/Copy of NSN Data/NSN Data (March 28 2023) - Point Count Observations.csv")

nsn.sampling <- read.csv("../Data_raw/Copy of NSN Data/NSN Data (March 28 2023) - Point Count Sampling.csv")

nsn.old <- read.csv("../Data_raw/Copy of NSN Data/NSN Data (March 28 2023) - Old Routes.csv")

#rt.id id numbers to fix
rt.id.correct <- read.csv("../Data_raw/NSN_routeid_corrections.csv",
                          colClasses = "character")
```

##Fix unmatched route id so matches routes (and location information)
```{r, echo=T}
#########
#routes

#convert 5 digit route id to 6 digit
nsn$Routeid.char <- sprintf("%06i", nsn$Routeid)
nsn$st_code <- substr(nsn$Routeid.char, 0, 2)
nsn$rt_code <- substr(nsn$Routeid.char, 3, 6)

# write.table(unique(nsn$st_code), file="../Data_raw/NSN_routeid_codes.txt", sep=",", row.names=F)
#determine states manually using graphing of points below

summary(nsn)
nsn.sf <- st_as_sf(nsn, coords=c("Latitude","Longitude"),
                   crs=projection(AF_states))
plot(nsn.sf)

# unique(nsn$Routeid)

# plot(nsn.sf[grep(nsn.sf$Routeid, pattern="^02"),])

base.plot <- ggplot() + geom_sf(data=nsn.sf, col="white") +
  geom_sf(data=AF_states) + theme_bw() +
  theme(panel.grid=element_blank())
# base.plot + geom_sf(data=nsn.sf[nsn.sf$st_code == "98",])

#read csv file back in
nsn.state.codes <- read.csv("../Data_raw/NSN_routeid_codes.csv",
                            colClasses = "character")

nsn <- merge(nsn, nsn.state.codes, by.x="st_code",
             by.y="routeid_code", all.x=T)

#########
#route id
summary(rt.id)
nrow(rt.id)
table(rt.id$state)

rt.id$route_id.char <- sprintf("%06i", rt.id$id)
rt.id$st_code <- substr(rt.id$route_id.char, 0,2)
rt.id$rt_code <- substr(rt.id$route_id.char, 3,6)

# verify <- unique(rt.id[,c("state","st_code")])
# write.table(verify[duplicated(verify$state),], sep=",", file = "../Data_raw/nsn_routeid_tofix.txt", row.names=F)

#read in corrected file and correct id numbers so they match nsn routes.csv
rt.id <- merge(rt.id.correct[,c("original","corrected")], rt.id,
               by.x="original", by.y="id", all.y=T)
rt.id[is.na(rt.id$corrected),]$corrected <- rt.id[is.na(rt.id$corrected),]$original

rt.id$route_id.char <- sprintf("%06i", as.integer(rt.id$corrected))
rt.id$st_code <- substr(rt.id$route_id.char, 0,2)
rt.id$rt_code <- substr(rt.id$route_id.char, 3,6)

unique(rt.id[,c("state","st_code")])

#NY, ME, MA, WI, NH, DE, VT, and AZ all have 00 state codes

######
#other corrections
rt.id[rt.id$st_code == 36 & rt.id$state == "ID",]$state <- "IA" #is this correct???

#check for duplicated rt.id$corrected
rt.id[duplicated(rt.id$corrected),]
nrow(rt.id)
rt.id <- rt.id[!duplicated(rt.id$corrected),]
nrow(rt.id) #removed 68 rows
```

##Try merging NSN data together to create master data frame

rt.id = 4325 rows
nsn = 3418 rows

```{r, echo=T}
# master <- merge(nsn, rt.id, all=T, by.x="Routeid", by.y="id",
#                 suffixes=c(".rt","rtid"))
#don't do this with original rt.id data frame since unmatched routeIDs

#this rt.id has location information already
master <- merge(rt.id, nsn, all=T,
                by.x="route_id.char", by.y="Routeid.char",
                suffixes=c("",".rt"))
#if has lat/long, then good, if not need to fix, add to NSN_rotueid_corrections.csv

  # master[is.na(master$Latitude) & master$state %in% AF_states$ST,]
  #data from Atlantic Flyway that doesn't have location information
  
  master[duplicated(master$corrected),]$route_id.char
  #"259249" "550037" "880001" #all of these have NA original & correct
  # master[master$route_id.char %in% c("259249"),]

master <- merge(master, nsn.time, by.x="corrected", by.y="route_id", 
                all=T, suffixes=c("",".time"))
nrow(master) #7374 rows
master <- merge(master, nsn.sampling, by.x="id",
                by.y="survey_id", all=T, suffixes=c("",".sampling"))
nrow(master) #49418 rows
master <- merge(master, nsn.obs, by.x="id.sampling",
                by.y="survey_stop_id", all=T, suffixes=c("",".obs"))
nrow(master)
master <- merge(master, nsn.sp, all.x=T, by.x="species_id", by.y="id",
                suffixes=c("",".sp"))
nrow(master)

# unique(master[is.na(master$id),]$route_id.char) #route sampling info and time/miles info, but no route info, no route id
#are these duplicated entries?
#221

#do this once
# write.table(master, file="../Data_processed/master_nsn_2023-06-21.txt",
#             sep=",", row.names=F)
```

## Raw data from non-Nightjar Survey Network
```{r, echo=T}

#read in raw data
non.nsn <- read.csv("../Data_raw/CombinedData/Allobs_noNJN.csv")
non.nsn.sites <- read.csv("../Data_raw/CombinedData/Allroute_info.csv")

missing.ny.sites <- read.

#read in CT data
# ct.1 <- read_excel("../Data_raw/CT Whippoorwill Data/WPWL2010.xls",
#                    sheet="DataSummary4")
#can also read in DataAnalysisInterval2 for counts by interval

ct.file.list <- dir("../Data_raw/CT Whippoorwill Data",
                    pattern=".xls", full.names = T)
ct.data.all <- data.frame()
for(f in ct.file.list)
{
  print(f)
  ct.data <- read_excel(f, sheet="DataSummary4")
  ct.data$file <- f
  ct.data$year <- as.numeric(gsub(gsub(f,
                  pattern="../Data_raw/CT Whippoorwill Data/WPWL",
                  replacement=""), pattern=".xls", replacement=""))
  ct.data.all <- rbind(ct.data.all, ct.data)
}
# write.table(file=
#               "../Data_raw/CT Whippoorwill Data/master_CT_WPWL.txt", 
#             ct.data.all, sep=",", row.names=F)

#fix incorrect 
non.nsn.sites[non.nsn.sites$Route.Longitude > 0 & !is.na(non.nsn.sites$Route.Longitude),]$Route.Longitude <- -(non.nsn.sites[non.nsn.sites$Route.Longitude > 0 & !is.na(non.nsn.sites$Route.Longitude),]$Route.Longitude)

table(non.nsn.sites$state)

# ontcan <- non.nsn.sites[non.nsn.sites$state == "ONTCAN",] #not in Atlantic flyway
# ma <- non.nsn.sites[non.nsn.sites$state == "MA",] 
# plot(ontcan$Route.Longitude, ontcan$Route.Latitude)
# plot(ma$Route.Longitude, ma$Route.Latitude) #no location information

unique(non.nsn.sites$state)

nrow(non.nsn.sites)

nrow(non.nsn.sites[is.na(non.nsn.sites$Route.Latitude),])
nrow(non.nsn.sites[non.nsn.sites$created_at == "",])

#some sites in the NE don't have created_at date times
# non.nsn.sites <- non.nsn.sites[!is.na(non.nsn.sites$Route.Latitude) & non.nsn.sites$created_at != "",] 
```

##Make master file of non-NSN sites and observations
NH sites: non.nsn.sites$id == non.nsn$route ish
  non.nsn.sites$id has - between state and site name
  
```{r, echo=T}
unique(non.nsn$state)

nh.sites <- non.nsn.sites[non.nsn.sites$state == "NH",]
  nh.sites$id <- gsub(nh.sites$id, pattern="-", replacement="")
nh.obs <- non.nsn[non.nsn$state == "NH",]
nh <- merge(nh.sites, nh.obs, all=T,
            by.x=c("id","Route.Stop","state"),
            by.y=c("route","stop","state"))
nh[is.na(nh$species),]$species <- "NONE"
nh[nh$species == "WHWP",]$species <- "EWPW"


ny.sites <- non.nsn.sites[non.nsn.sites$state == "NY",]
ny.missing.sites <- read.csv("../Data_raw/CombinedData/NY_MissingLocations_fromMattPalumbo.csv")
  #only need to do below once to convert NAD83 zone 18 to lat/long (copy and pasted into csv file)
  # ftdrum <- ny.missing.sites[ny.missing.sites$route == "Fort Drum",]
  # ftdrum.sf <- st_as_sf(ftdrum, coords=c("N","W"), crs=26918)
  # ftdrum.sf.latlong <- st_transform(ftdrum.sf, crs=crs(AF_states))
ny.missing.sites <- ny.missing.sites[!is.na(ny.missing.sites$N),]

ny.obs <- non.nsn[non.nsn$state == "NY",]
  #need to correct route information so it matches ny.sites
  ny.correct <- read.csv("../Data_raw/combined_corrections.csv")
  ny.obs <- merge(ny.obs, ny.correct[,c("route","correct")],
                  by="route", all.x=T)
ny <- merge(ny.sites, ny.obs, all=T,
            by.x=c("name","Route.Stop","state"),
            by.y=c("correct","stop","state"))
ny[is.na(ny$species),]$species <- "NONE"
ny[ny$species == "-",]$species <- "NONE"
ny$species <- toupper(ny$species)
  #need to correct species codes to match other data files
  ny.sp <- read.csv("../Data_raw/combined_sp_key.csv")
  ny <- merge(ny.sp[,c("original_code","code")], ny, 
              by.y="species", by.x="original_code", all.y=T)
  names(ny) <- c("original_code","species", names(ny)[3:25])
  
vt.sites <- non.nsn.sites[non.nsn.sites$state == "VT",]
vt.obs <- non.nsn[non.nsn$state == "VT",] #no Vermont observations received as of 6/26/2023

vt <- merge(vt.sites, vt.obs, all=T,
            by.x=c("name","Route.Stop","state"),
            by.y=c("route","stop","state"))

names(nh)
# "id"              "Route.Stop"      "state"           "name"
# "created_at"      "updated_at"     "Route.Latitude""Route.Longitude"
# "year"            "stop_des"        "wind"            "sky" 
# "noise"           "cars"            "moon"            "dir"           "species"         "min.1"           "min.2"           "min.3"         "min.4"           "min.5"           "min.6"          

#match columns with ny, nh, and vt data
ct.data.all$stop <- as.numeric(substr(ct.data.all$PointID,
                           start=nchar(ct.data.all$PointID)-1, 
                           stop = nchar(ct.data.all$PointID)))
ct.data.all$species <- "NONE"
ct.data.all[ct.data.all$MaxCount > 0,]$species <- "EWPW"
ct.data.all$MaxCount <- as.numeric(ct.data.all$MaxCount)

ct <- data.frame(ct.data.all[,c("PointID","stop","RouteName",
                 "Latitude","Longitude","year","species","SumCount",
                 "MaxCount")])
names(ct) <- c("id","Route.Stop","name",
               "Route.Latitude","Route.Longitude",
               "year","species","TotalDetections","MaxDetections")
ct$state <- "CT"
ct$Route.Longitude <- -ct$Route.Longitude

non.nsn.master <- rbind(nh, ny[,names(nh)], vt[,names(nh)])
non.nsn.master$TotalDetections <-
  rowSums(non.nsn.master[,grep(names(non.nsn.master),
                               pattern="min.")])
non.nsn.master$MaxDetections <- 
  apply(non.nsn.master[,grep(names(non.nsn.master), pattern="min.")],
        1, function(x) max(x, na.rm=T))
non.nsn.master[non.nsn.master$MaxDetections == "-Inf",]$MaxDetections <- NA
non.nsn.master <- bind_rows(non.nsn.master, ct)

#only run once
# write.table(non.nsn.master,
#             file="../Data_processed/master_nonnsn_2023-06_21.txt",
#             sep=",", row.names=F)
```

```{r, echo=F}
##Map non NSN sites in Atlantic flyway
#not relevant anymore. only observations were in NY and NH in obs file
# #subset non.nsn.sites to only Atlantic flyway sites
# non.nsn.sites <- non.nsn.sites[non.nsn.sites$state %in%
#                                  unique(AF_states$ST),]
# 
# non.nsn.sites.sf <- st_as_sf(non.nsn.sites[!is.na(non.nsn.sites$Route.Latitude),], coords=c("Route.Longitude","Route.Latitude"), crs=projection(AF_states))
# 
# base.plot <- ggplot() + geom_sf(data=nsn.sf, col="white") + geom_sf(data=AF_states) + theme_bw() + theme(panel.grid=element_blank())
# base.plot + geom_sf(data=non.nsn.sites.sf)
# 
# af.plot <- ggplot() + geom_sf(data=AF_states, fill="white") + theme_bw() + theme(panel.grid=element_blank())
# af.plot
# af.plot + geom_sf(data=non.nsn.sites.sf)
```